---
title: "LEMMA Model Descriptive"
author: "Chris Hoover"
date: "5/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)

```

## Equations  
\begin{align*}
\dot{S}&=-\beta S(I_R+I_H)/N \\
\dot{E}&=\beta S(I_R+I_H)/N-\sigma E \\
\dot{I_R}&=\sigma(1-\alpha) E-\gamma_R I_R \\
\dot{I_H}&=\sigma\alpha E-\gamma_H I_H \\
\dot{H}&=\gamma_H I_H-\psi H \\
\dot{DC}&=\psi H \\
\dot{R}&=\gamma_R I_R \\
\end{align*}

```{r par_table, echo = FALSE, include = FALSE, dev="tikz"}
tab1 <- data.frame(des = c("Susceptibles",
                           "Exposed, pre-infectious",
                           "Infectious, will be hospitalized",
                           "Infectious, will no be hospitalized",
                           "Hospitalized",
                           "Discharged from hospital (either deceased or recovered)",
                           "Recovered without being hospitalized",
                           "Total population size",
                           "Transmission rate",
                           "1/latent period",
                           "Proportion infected that will be hospitalized",
                           "1/time from infectiousness to hospitalization",
                           "1/duration of infectiousness for non-hospital-bound cases",
                           "1/time from hospitalization to discharge"
))

rownames(tab1) <- c("$S$","$E$","$I_H$", "$I_R$","$H$","$DC$", "$R$", "$N$",
                    "$\\beta$","$\\sigma$", "$\\alpha$","$\\gamma_H$", "$\\gamma_R$", "$\\psi$")

knitr::kable(tab1, row.names = TRUE, 
             #col.names = c("Value", "Definition"), 
             format = "latex", escape = FALSE, 
             caption = "Parameter values and descriptions used in the model")
```


## $\mathcal{R}_0$ formulation  
For complex models such as this, best to use the next generation matrix method (see [here](https://doi.org/10.1098/rsif.2009.0386)) to estimate the basic reproduction number as a function of model equations. Briefly, the next generation matrix method constructs transmission, $\mathbf{T}$, and transition, $\mathbf{\Sigma}$, matrices from the linearized infection subsystem of a system of differential equations and $\mathcal{R}_0$ is the max eigenvalue of $-\mathbf{T\Sigma}^{-1}$. User can select whether hospital population, $H$, is infectious(and therefore part of the infection subsystem) or not, so will construct estimates for case where $H$ is and is not infectious.

### Hospital population, $H$, is infectious  
$$\mathbf{T}=
\begin{bmatrix}
0 & \beta & \beta & \beta \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
\end{bmatrix}
\mathbf{\Sigma}=
\begin{bmatrix}
-\sigma & 0 & 0 & 0 \\
\sigma(1-\alpha) & \gamma_R & 0 & 0 \\
\sigma\alpha & 0 & -\gamma_H & 0 \\
0 & 0 & \gamma_H & -\psi \\
\end{bmatrix}$$


$$\mathcal{R}_0=\beta\sigma\frac{(1-\alpha)\gamma_H+\alpha\gamma_R}{\gamma_R\gamma_H\psi}$$

### Hospital population, $H$, is NOT infectious  
$$\mathbf{T}=
\begin{bmatrix}
0 & \beta & \beta \\
0 & 0 & 0  \\
0 & 0 & 0  \\
\end{bmatrix}
\mathbf{\Sigma}=
\begin{bmatrix}
-\sigma & 0 & 0  \\
\sigma(1-\alpha) & \gamma_R & 0  \\
\sigma\alpha & 0 & -\gamma_H  \\
\end{bmatrix}$$


$$\mathcal{R}_0=\beta\sigma\frac{(1-\alpha)\gamma_H+\alpha\gamma_R}{\gamma_R\gamma_H}$$