---
title: "Stan with PUI"
author: "Chris Hoover"
date: "5/25/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE)

require(tidyverse)
require(rstan)

# Rstan prep
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')

```

# Purpose  
Compare model fits when incorporating data on persons under investigation (PUI) in hospitals. PUIs are those who are suspected as potential COVID cases who have not yet been confirmed positive.

The STAN model was slightly edited to include PUI by added functionality similar to how the fraction of infected cases which are detected is implemented. An additional prior is incorporated to represent the proportion of PUI who are positive. The model then incorporates this prior times the observed number of PUIs when evaluating model fit (e.g. rather than taking hospitalized and icu as observed, takes hospitalized+frac_pui_pos*hospitalized_puis as observed).

# Data   
Need two files: one with cases that has columns `Date`, `cum_cases`, and `cum_death` corresponding to the date of observation and the cumulative number of COVID cases and deaths confirmed on that date. The other with columns `date`, `ICU_PUI`, `ICU_CONF`, `HOSP_PUI`, and `HOSP_CONF` corresponding to the date of observation, the number of people under investigation in the ICU, the number of confirmed COVID patients in the ICU, and the same for hospitalized COVID patients (PUI and confirmed patients).

Also takes a pretty hefty parameters list with demographics, priors, and some model settings.

```{r get_data}
# Case and hosptialization data for SF
sf_case <- read.csv(url("https://data.sfgov.org/resource/tvq9-ec9w.csv")) %>% 
  mutate(Date = as.Date(date)) %>% 
  pivot_wider(names_from = case_disposition,
              values_from = case_count) %>% 
  group_by(Date) %>% 
  summarise(Cases = sum(Confirmed, na.rm = T),
            Deaths = sum(Death, na.rm = T)) %>% 
  arrange(Date) %>% 
  mutate(cum_case = cumsum(Cases),
         cum_death = cumsum(Deaths))

sf_hosp <- read.csv(url("https://data.sfgov.org/resource/nxjg-bhem.csv")) %>% 
  mutate(Date = as.Date(reportdate),
         type = ifelse(dphcategory == "ICU", "ICU", "HOSP"),
         conf = ifelse(covidstatus == "PUI", "PUI", "CONF"),
         hosp_stat = paste(type, conf, sep = "_")) %>% 
  pivot_wider(names_from = hosp_stat,
              values_from = patientcount) %>%
  group_by(Date) %>% 
  summarise(ICU_PUI = sum(ICU_PUI, na.rm = T),
            ICU_CONF = sum(ICU_CONF, na.rm = T),
            HOSP_PUI = sum(HOSP_PUI, na.rm = T),
            HOSP_CONF = sum(HOSP_CONF, na.rm = T)) %>% 
  arrange(Date)

saveRDS(sf_case, 
        file = paste("Data/sf_case_data_", Sys.Date(), ".RDS", sep = ""))

saveRDS(sf_hosp, 
        file = paste("Data/sf_hosp_data_", Sys.Date(), ".RDS", sep = ""))
```

```{r parameters}
start.date = "2020-02-16"
end.date = "2020-07-01"

seir_inputs <- list()

# the number of age groups
  seir_inputs[['nage']] = 1 

# the population for each age group
  seir_inputs[['npop']] = array(883305)
# the fraction of hospitalized cases (ICU + non-ICU)
  seir_inputs[['frac_hosp']] = array(0.07)
# the fraction of ICU cases
  seir_inputs[['frac_icu']] = array(0.02)
# the death rate of infected
  seir_inputs[['frac_mort']] = array(0.01)
# the fraction of asymptomatic cases
  seir_inputs[['frac_asym']] = array(0.178)

# a parameter modifying the prior uncertainty in the fractions above (set to 1000.0 for `nage` > 1)
  seir_inputs[['alpha_multiplier']] = 100.0

# mean duration in "exposed" stage
  seir_inputs[['mu_duration_lat']] = 5.0
# standard deviation (sd) of duration in "exposed" stage
  seir_inputs[['sigma_duration_lat']] = 2.0
# mean duration in "infectious" stage for asymptomatic cases
  seir_inputs[['mu_duration_rec_asym']] = 7.0
# sd of duration in "infectious" stage for asymptomatic cases
  seir_inputs[['sigma_duration_rec_asym']] = 5.0
# mean duration in "infectious" stage for mild cases
  seir_inputs[['mu_duration_rec_mild']] = 7.0
# sd of duration in "infectious" stage for mild cases
  seir_inputs[['sigma_duration_rec_mild']] = 5.0
# mean duration in "infectious" stage for hospitalized cases
  seir_inputs[['mu_duration_pre_hosp']] = 5.0
# sd of duration in "infectious" stage for hospitalized cases
  seir_inputs[['sigma_duration_pre_hosp']] = 1.0
# mean duration in hospital for non-ICU cases
  seir_inputs[['mu_duration_hosp_mod']] = 7.0
# sd of duration in hospital for non-ICU cases
  seir_inputs[['sigma_duration_hosp_mod']] = 1.0
# mean duration in hospital for ICU cases
  seir_inputs[['mu_duration_hosp_icu']] = 16.0
# sd of duration in hospital for ICU cases
  seir_inputs[['sigma_duration_hosp_icu']] = 4.0

# mean fraction of tested infectious
  seir_inputs[['mu_frac_tested']] = 0.25
# sd fraction of tested infectious
  seir_inputs[['sigma_frac_tested']] = 0.1
  
# lambda parameter for initial conditions of "exposed"
  seir_inputs[['lambda_ini_exposed']] = 0.3

# mean initial beta estimate
  seir_inputs[['mu_beta1']] = 0.2
# sd initial beta estimate
  seir_inputs[['sigma_beta1']] = 0.02

# distance in time between the knots used to construct the splines 
  seir_inputs[['dknot']] = 10

# spline mode (must be 1 or 2)
# splinemode = 1: estimate beta up to today
# splinemode = 2: estimate beta up to dknot days before today, then assume constant value up to today
  seir_inputs[['splinemode']] = 1

# number of interventions
  seir_inputs[['ninter']] = 1
  
# length of each intervention
  seir_inputs[['len_inter']] = array(10)
# mean change in beta through intervention 
  seir_inputs[['mu_beta_inter']] = array(1.0)
# sd change in beta through intervention
  seir_inputs[['sigma_beta_inter']] = array(0.5)

```

# Run model without PUI (done silently)  
```{r stan_original, eval = FALSE}
case_dat = sf_case
hosp_dat = sf_hosp

# the start date at which to start the simulation
  startdate_str = as.Date(start.date)

# Timeframe for simulation. Can replace Sys.Date() with any date at which to cut off model fitting
  itoday_1based = Sys.Date()-startdate_str
  
# specify the date up to when beta is estimated
  seir_inputs[['itoday']] = as.numeric(itoday_1based)

# the number of days to run the model for
  seir_inputs[['nt']] = as.numeric(as.Date(end.date) - as.Date(start.date))
  
# Put data into list
  seir_inputs[['nobs']] <- nrow(case_dat)
  seir_inputs[['tobs']] <- as.numeric(case_dat$Date - startdate_str)
  seir_inputs[['obs_I']] <- case_dat$cum_case
  seir_inputs[['obs_Rmort']] <- case_dat$cum_death
  
  seir_inputs[['nhobs']] <- nrow(hosp_dat)
  seir_inputs[['thobs']] <- as.numeric(hosp_dat$Date - startdate_str)
  seir_inputs[['obs_Hmod']] <- hosp_dat$HOSP_CONF
  seir_inputs[['obs_Hicu']] <- hosp_dat$ICU_CONF
  
# Note: The inputs below must be lists or arrays, even if one intervention is specified 
# (the enclosing "[]" are not optional). For zero interventions 

# start time of each interventions
  seir_inputs[['t_inter']] = array(as.numeric(itoday_1based+5))
  
# NOTE: all other parameterss should already be in pars  
  
# Run the stan model  
  stan_mod <- "UCSC_SEIR.stan"
  
stan_run <- stan(
  file = stan_mod,
  data = seir_inputs,
  thin = 2,
  chains = 3,
  warmup = 500,
  iter = 2500,
  cores = 3,
  refresh = 100
)

stan_no_pui <- extract(stan_run, permuted = T)

saveRDS(stan_no_pui,
        file = paste("Outputs/", str_split(stan_mod, ".stan")[[1]][1], "_", "SF_", Sys.Date(), ".RDS", sep = ""))

```

# Run model with PUI (done silently)  
```{r stan_pui, eval = FALSE}
seir_inputs_pui <- seir_inputs

# Observed PUI data
  seir_inputs_pui[['obs_Hpui']] <- hosp_dat$HOSP_PUI
  seir_inputs_pui[['obs_Hpui_icu']] <- hosp_dat$ICU_PUI

# mean fraction of pui who end up positive
  seir_inputs_pui[['mu_pui_pos']] = 0.25
# sd fraction of pui who end up positive
  seir_inputs_pui[['sigma_pui_pos']] = 0.1

stan_mod_pui <- "SEIR_PUI.stan"
  
stan_run2 <- stan(
  file = stan_mod_pui,
  data = seir_inputs_pui,
  thin = 2,
  chains = 3,
  warmup = 500,
  iter = 2500,
  cores = 3,
  refresh = 100
)

stan_pui <- extract(stan_run2, permuted = T)

saveRDS(stan_pui,
        file = paste("Outputs/", str_split(stan_mod_pui, ".stan")[[1]][1], "_", "SF_", Sys.Date(), ".RDS", sep = ""))
  
```

# Compare model fits  
```{r sum_fits}
stan_no_pui <- readRDS("Outputs/UCSC_SEIR_SF_2020-05-25.RDS")
stan_pui <- readRDS("Outputs/SEIR_PUI_SF_2020-05-25.RDS")

# Summary function to get summaries across posteriors
stan_post_sum <- function(d){
  thing.med <- median(d)
  thing.mean <- mean(d)
  thing.75 <- quantile(d, 0.75)
  thing.25 <- quantile(d, 0.25)
  thing.975 <- quantile(d, 0.975)
  thing.025 <- quantile(d, 0.025)
  
  return(c("q.mean" = thing.mean, 
           "q.50" = thing.med, 
           "q.75" = thing.75, 
           "q.25" = thing.25, 
           "q.975" = thing.975, 
           "q.025" = thing.025))
}

# State variables summary from model
state_vars <- c("S", "E", "Ia", "Im", "Ip", "Hosp", "Hicu", "Rlive", "Rdead")
  
# No pui run summary
no_pui_state_vars <- do.call(rbind, lapply(1:dim(stan_no_pui$x)[2], function(s){
  as_tibble(t(apply(stan_no_pui$x[,s,],2,stan_post_sum))) %>% 
    mutate(State = state_vars[s],
           obs = row_number(),
           Date = as.Date(start.date)+obs)
}))

colnames(no_pui_state_vars)[1:6] <- c("var_mean", "var_med", "var_75th", "var_25th", "var_975th", "var_025th")

# Observed infections (since we don't observe all infections output by actual model)
I_obs_no_pui <- stan_no_pui[["mu_Iobs"]]

I_obs_no_pui <- t(apply(I_obs_no_pui, 2, stan_post_sum))

colnames(I_obs_no_pui) <- c("Iobs_mean", "Iobs_med", "Iobs_75th", "Iobs_25th", "Iobs_975th", "Iobs_025th")
 
I_obs_no_pui <- I_obs_no_pui %>% as_tibble() %>% 
  mutate(obs = row_number(),
         Date = as.Date(start.date)+obs)


# PUI run  summary
pui_state_vars <- do.call(rbind, lapply(1:dim(stan_pui$x)[2], function(s){
  as_tibble(t(apply(stan_pui$x[,s,],2,stan_post_sum))) %>% 
    mutate(State = state_vars[s],
           obs = row_number(),
           Date = as.Date(start.date)+obs)
}))

colnames(pui_state_vars)[1:6] <- c("var_mean", "var_med", "var_75th", "var_25th", "var_975th", "var_025th")

# Observed infections (since we don't observe all infections output by actual model)
I_obs_pui <- stan_pui[["mu_Iobs"]]

I_obs_pui <- t(apply(I_obs_pui, 2, stan_post_sum))

colnames(I_obs_pui) <- c("Iobs_mean", "Iobs_med", "Iobs_75th", "Iobs_25th", "Iobs_975th", "Iobs_025th")
 
I_obs_pui <- I_obs_pui %>% as_tibble() %>% 
  mutate(obs = row_number(),
         Date = as.Date(start.date)+obs)

```


## Posterior distribution of pui positive from pui model  
```{r pui_pos_dist}
hist(stan_pui$frac_puipos, breaks = 30,
     xlab = "Fraction PUI positive")
```

## Case numbers  
```{r comp_case_nums}
pui_case_comp <- rbind(I_obs_no_pui %>% 
                         mutate(PUI = "No"),
                       I_obs_pui %>% 
                         mutate(PUI = "YES"))

ggplot() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11, 
                                   angle = 45,
                                   hjust = 1),
        axis.title = element_text(size = 14)) +
  geom_ribbon(data = pui_case_comp,
              aes(x = Date, 
                  ymax = Iobs_75th,
                  ymin = Iobs_25th,
                  fill = PUI),
              alpha = 0.4) +
  geom_ribbon(data = pui_case_comp,
              aes(x = Date, 
                  ymax = Iobs_975th,
                  ymin = Iobs_025th,
                  fill = PUI),
              alpha = 0.2) +
  geom_point(data = sf_case,
             aes(x = Date, y = cum_case),
             pch = 17) +
  geom_line(data = pui_case_comp,
            aes(x = Date, y = Iobs_mean, col = PUI)) +
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = as.Date(c("2020-02-16",
                                  "2020-07-04")))+
  ylim(c(0,5000)) +
  labs(y = "cumulative infections",
       title = "SF Observed and fit infections")

```
Not much of a difference in the fit, but projections begin to diverge after awhile.

## Hospitalizations  
```{r comp_hosp_numbers}
pui_hosp_comp <- rbind(pui_state_vars %>% 
                         filter(State == "Hosp") %>% 
                         mutate(PUI = "No"),
                       no_pui_state_vars %>% 
                         filter(State == "Hosp") %>% 
                         mutate(PUI = "YES"))


ggplot() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11, 
                                   angle = 45,
                                   hjust = 1),
        axis.title = element_text(size = 14)) +
# Observed data
  geom_point(data = sf_hosp,
             aes(x = Date, y = HOSP_CONF),
             pch = 17) +
  geom_point(data = sf_hosp,
             aes(x = Date, y = HOSP_CONF+HOSP_PUI),
             pch = 16) +
  geom_segment(data = sf_hosp,
               aes(x = Date, 
                   xend = Date,
                   y = HOSP_CONF,
                   yend = HOSP_CONF + HOSP_PUI)) +
# Hospitalizations uncertainty intervals  
  geom_ribbon(data = pui_hosp_comp,
              aes(x = Date, 
                  ymax = var_75th,
                  ymin = var_25th,
                  fill = PUI),
              alpha = 0.4) +
  geom_ribbon(data = pui_hosp_comp,
              aes(x = Date, 
                  ymax = var_975th,
                  ymin = var_025th,
                  fill = PUI),
              alpha = 0.2) +
# Model means
  geom_line(data = pui_hosp_comp,
            aes(x = Date, y = var_mean, col = PUI)) +
# Formatting
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = as.Date(c("2020-02-16",
                                  "2020-06-20"))) +
  ylim(c(0,max(sf_hosp$HOSP_PUI+sf_hosp$HOSP_CONF, na.rm = T)+25)) +
  labs(y = "Hospitalizations",
       title = "Comparison btwn fits including PUI or not")

```

Not much of a difference here which I guess makes sense since we're not really changing any of the fitting criteria. Would have to allow PUI to influence how the model fits in order to start really seeing differences, I think, but that would require incorporating some mechanism to generate PUI in the model which seems like more work than it's worth.

## ICUs  
```{r comp_icu_numbers}
pui_icu_comp <- rbind(pui_state_vars %>% 
                         filter(State == "Hicu") %>% 
                         mutate(PUI = "No"),
                       no_pui_state_vars %>% 
                         filter(State == "Hicu") %>% 
                         mutate(PUI = "YES"))


ggplot() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11, 
                                   angle = 45,
                                   hjust = 1),
        axis.title = element_text(size = 14)) +
# Observed data
  geom_point(data = sf_hosp,
             aes(x = Date, y = ICU_CONF),
             pch = 17) +
  geom_point(data = sf_hosp,
             aes(x = Date, y = ICU_CONF+ICU_PUI),
             pch = 16) +
  geom_segment(data = sf_hosp,
               aes(x = Date, 
                   xend = Date,
                   y = ICU_CONF,
                   yend = ICU_CONF + ICU_PUI)) +
# Hospitalizations uncertainty intervals  
  geom_ribbon(data = pui_icu_comp,
              aes(x = Date, 
                  ymax = var_75th,
                  ymin = var_25th,
                  fill = PUI),
              alpha = 0.4) +
  geom_ribbon(data = pui_icu_comp,
              aes(x = Date, 
                  ymax = var_975th,
                  ymin = var_025th,
                  fill = PUI),
              alpha = 0.2) +
# Model means
  geom_line(data = pui_icu_comp,
            aes(x = Date, y = var_mean, col = PUI)) +
# Formatting
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = as.Date(c("2020-02-16",
                                  "2020-06-20"))) +
  ylim(c(0,max(sf_hosp$ICU_PUI+sf_hosp$ICU_CONF, na.rm = T)+25)) +
  labs(y = "ICU hospitalizations",
       title = "Comparison btwn fits including PUI or not")

```
