---
title: "ABM Testing"
author: "Chris Hoover"
date: "6/15/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

require(tidyverse)
```

## Recap  
Can fit a simple model to data we trust (hospitalizations and/or deaths) and use the resulting simulations to determine the degree of underreporting from testing data if we take the observed testing data as our sample and our model simulations as the truth. can then project into the future with different assumptions about the number of tests available and future transmission conditions (e.g. $\mathcal{R}_e$) to get an idea of what kind of observations in testing data might be cause for concern (e.g. serve as triggers for return to SiP or other interventions) and how these observations/triggers compare with other potential triggers (e.g. hospital admissions).

This analysis is a good starting point, but testing does not influence infection in hte model (maybe this is negligible, but maybe not) and the assumption that the fitted model accurately represents true infection dynamics is quite a stretch. So, building on this analysis with an ABM with testing incorporated in order to get a better idea of how sampling weights (e.g. test-seeking behavior) between infection classes influences the observed number of positive tests, how knowledge/reporting of this heterogeneity may give us better estimates of key metrics like $\mathcal{R}_e$ and incidence, and how we might better allocate testing resources to estimate these metrics.

```{r epi_data}
# Case and hosptialization data for SF
sf_hosp <- read.csv(url("https://data.sfgov.org/resource/nxjg-bhem.csv")) %>% 
  mutate(Date = as.Date(reportdate),
         type = ifelse(dphcategory == "ICU", "ICU", "HOSP"),
         conf = ifelse(covidstatus == "PUI", "PUI", "CONF"),
         hosp_stat = paste(type, conf, sep = "_")) %>% 
  pivot_wider(names_from = hosp_stat,
              values_from = patientcount) %>%
  group_by(Date) %>% 
  summarise(ICU_PUI = sum(ICU_PUI, na.rm = T),
            ICU_CONF = sum(ICU_CONF, na.rm = T),
            HOSP_PUI = sum(HOSP_PUI, na.rm = T),
            HOSP_CONF = sum(HOSP_CONF, na.rm = T)) %>% 
  arrange(Date) %>% 
  mutate(HOSP_tot = ICU_CONF + HOSP_CONF,
         HOSP_max = ICU_CONF + HOSP_CONF + ICU_PUI + HOSP_PUI,
         cumICUconf = cumsum(ICU_CONF),
         cumICUpui = cumsum(ICU_PUI),
         cumHOSPconf = cumsum(HOSP_CONF),
         cumHOSPpui = cumsum(HOSP_PUI))

sf_case <- read.csv(url("https://data.sfgov.org/resource/tvq9-ec9w.csv")) %>% 
  mutate(Date = as.Date(specimen_collection_date)) %>% 
  pivot_wider(names_from = case_disposition,
              values_from = case_count) %>% 
  group_by(Date) %>% 
  summarise(Cases = sum(Confirmed, na.rm = T),
            Deaths = sum(Death, na.rm = T)) %>% 
  arrange(Date) %>% 
  mutate(cum_case = cumsum(Cases),
         cum_death = cumsum(Deaths))

sf_test <- read.csv(url("https://data.sfgov.org/resource/nfpa-mg4g.csv")) %>% 
  mutate(Date = as.Date(specimen_collection_date)) %>% 
  arrange(Date) %>% 
  mutate(cum_tests = cumsum(tests),
         cum_pos = cumsum(pos))

sf_all <- sf_test %>% 
  dplyr::select(Date, tests, pos, neg, pct, indeterminate, cum_tests, cum_pos) %>%
  full_join(sf_case, by = "Date") %>% 
  full_join(sf_hosp, by = "Date")
```

# ABM Simulations with Testing built in  
```{r abm_setup, include = FALSE}
devtools::load_all("../ABM/")

set.seed(430)

# Number of people, key dates, transmission probability 
# Key Dates
t0 <- as.Date("2020-02-15")
last_dat <- max(sf_all %>% na.omit() %>% pull(Date))
today <- Sys.Date()

# Beta change dates
db1 <- as.Date("2020-03-05") #Schools closed
db2 <- as.Date("2020-03-15")     #SiP announced
db3 <- as.Date("2020-04-15")          #Another knot
db4 <- as.Date("2020-05-15")          #Another knot
db5 <- as.Date("2020-06-15")         #Another knot 
t.end <- as.Date("2020-08-01")         # Where we're headed (for future use)

t.tot <- as.numeric(t.end - t0)

  pop.red <- 10
  N <- 9e5/pop.red
  dt = 1/3
  
  beta.init <- 0.625 # Corresponds to R0 of ~3
  
  beta.time <- cbind(-as.numeric(t0-c(t0, db1,
                                      db2, db3, db4, db5, t.end)),
                     beta.init*c(1, 1, 0.6, 0.25, 0.3, 0.5, 0.6))
  
beta_smooth <- splinefun(x = beta.time[,1],
                         y = beta.time[,2],
                         method = "natural")

abm_beta_fx <- approxfun(c(0:t.tot),
                         c(rep(beta.init, beta.time[2,1]), sapply(c(beta.time[2,1]+1):(t.tot+1),
                                                                  beta_smooth)))
  
  as.data.frame(cbind(t = c(0:t.tot),
                      beta = sapply(0:t.tot, abm_beta_fx))*(1/beta.init)) %>% 
    mutate(Date = as.Date(t0+t)) %>% 
    ggplot(aes(x = Date, y = beta))+
      theme_bw() +
      geom_line() +
      labs(x = "Date", y = "Relative beta")
  
# Tests conducted
  test_fx <- approxfun(0:t.tot,
                       c(rep(0, as.numeric(min(sf_test$Date)-t0)+1),
                         sf_test$tests,
                         rep(2000, as.numeric(t.end-max(sf_test$Date))))*dt/pop.red,
                       method = "linear")
  
#Testing probabilities of different infection classes
  prob_test <- c("S" = 0.01,
                 "E" = 0.01,
                 "Ip" = 0.01,
                 "Ia" = 0.01,
                 "Im" = 0.1,
                 "Imh" = 0.1,
                 "Ih" = 1,
                 "R" = 0.01)

# Age distributions from US census  
  pop.props <- c(0.06+0.061, 
                 0.065+0.066, 
                 0.066+0.5*0.139, 
                 0.5*0.139+0.5*0.127, 
                 0.127, 0.5*0.127+0.066, 
                 0.063+0.5*0.093, 
                 0.5*0.093+0.5*0.048,
                 0.5*0.048+0.019)
  
  names(pop.props) <- c(seq(0,70,10), 80)  # 80 is anyone older than 80
  pop.ages <- as.numeric(sample(names(pop.props), size = N/pop.red, replace = TRUE, prob = pop.props))
  
# Baseline risk of infection based on age
  # This gives decent approximation of more risk among age groups that are more active, but could use improvement
  risks <- rgamma(N/pop.red, 3, 3)*(pop.ages+1)^(1/pop.ages)
  
  #hist(risks, breaks = 30, main = "Distribution of individual variability in risk")
  
# Quarantine probabilities 
  q.probs <- rbeta(N/pop.red, 5, 1)
  
  #hist(q.probs, breaks = 30, main = "Distribution of individual probability of quarantining given symptoms")

# Infection status through time
  inf.mat <- matrix(data = NA, nrow = N/pop.red, ncol = t.tot/dt)

# Test status through time
  test.mat <- matrix(data = NA, nrow = N/pop.red, ncol = t.tot/dt)
  
# Event times
  t.til.nxt <- matrix(data = NA, nrow = N/pop.red, ncol = t.tot/dt)
  

# Initial conditions
  e.seed <- 5     #Exposed
  ip.seed <- 1    #infected pre-symptomatic
  ia.seed <- 0    #infected asymptomatic
  im.seed <- 0    #infected mildly symptomatic
  imh.seed <- 0   #infected mildly symptomatic, will become severe
  ih.seed <- 0    #infected severely symptomatic
  d.seed <- 0     #dead
  r.seed <- 0     #removed
  s.seed <- N/pop.red - e.seed - ip.seed - ia.seed - im.seed - imh.seed - ih.seed - d.seed - r.seed

# Initialize infection and waiting time matrices    
init.infection <- sample(c(rep("E", e.seed),
                           rep("Ip", ip.seed),
                           rep("Ia", ia.seed),
                           rep("Im", im.seed),
                           rep("Imh", imh.seed),
                           rep("Ih", ih.seed),
                           rep("D", r.seed),
                           rep("R", r.seed),
                           rep("S", s.seed)), N/pop.red, replace = FALSE)
  
# Fill infection and state transition matrices based on initial conditions  
  inf.mat[,1] <- init.infection  
  t.til.nxt[,1] <- sapply(inf.mat[,1], function(i){
    if(i == "E"){
      t_latent()
    } else if(i == "Ip"){
      t_presymp()
    } else if(i == "Ia"){
      t_asymp()
    } else if(i == "Im"){
      t_msymp()
    } else if(i == "Imh"){
      t_mtosev()
    } else if(i == "Ih"){
      t_sevsymp()
    } else {
      NA
    }
  })
  
```

```{r abm_more_fxs}
# Simulate infection across population  
sim.infection <- function(inf.vec, q.vec, risks, bta){
  infecteds <- which(inf.vec %in% c("Ip", "Ia", "Im", "Imh") &
                       q.vec == 0)
  
  I.tot <- length(infecteds)
  inf.probs <- 1-exp(-bta*I.tot*risks)
  S.vec <- which(inf.vec == "S")
  StoEs <- sapply(S.vec, function(p){
    rbinom(1,1,inf.probs[p])
  })
  return(S.vec[which(StoEs == 1)])
}  

```

```{r abm_sim, eval = FALSE}
for(t in 2:(t.tot/dt)){
  # print(t)
# Advance transition times
    t.til.nxt[,t] <- t.til.nxt[,(t-1)]-dt
  
  # Advance expired states to next state 
    nexts <- which(t.til.nxt[,t] < 0) # states that expired
    if(length(nexts >0)){
      next.mat <- t(mapply(next_state, inf.mat[nexts,(t-1)], pop.ages[nexts])) # function returns new states and time spent in new state
      inf.mat[nexts,t] <- next.mat[,1] # update infection status matrix
      t.til.nxt[nexts,t] <- as.numeric(next.mat[,2]) # update waiting time matrix
    }
    
  # Anyone who hasn't changed infection state remains the same
    sames <- which(is.na(inf.mat[,t]))
    inf.mat[sames,t] <- inf.mat[sames,(t-1)]

  # Implement testing
    n_tests <- round(test_fx(t*dt))
    test.mat[,t] <- test_folks(inf.mat[,t], test.mat[,(t-1)], prob_test, n_tests)
    
  # Implement quarantining  
    quars <- quarantine(inf.mat[,t], test.mat[,t], q.probs)
    
  # Simulate infection
    beta_t <- abm_beta_fx(t*dt)  
    new.Es <- sim.infection(inf.mat[,t], quars, risks, beta_t)

  # Update infection status matrix with new Es  
    inf.mat[new.Es,t] <- "E"
    
  # Update transition times matrix for new Es from sample of latent period dist'n
    t.til.nxt[new.Es,t] <- as.numeric(replicate(length(new.Es), t_latent()))
  
  # On to the next one  
}

pos_tests <- t(apply(test.mat,2,function(x){
  Ttot = sum(!is.na(x), na.rm = T) ; return(Ttot)
}))[1,]

tests = round(test_fx(1:(t.tot/dt)*dt))

inf_sum <- as.data.frame(sum_inf_mat(inf.mat)) %>%
  mutate(t = (t.sim*dt),
         date = as.character(t0+t),
         tests = tests,
         tot_pos = pos_tests,
         pos = tot_pos-lag(tot_pos)) %>% 
  group_by(date) %>% 
  summarise(S = mean(S),
            E = mean(E),
            Ip = mean(Ip),
            Ia =mean(Ia),
            Im = mean(Im),
            Imh = mean(Imh),
            Ih = mean(Ih),
            R = mean(R),
            D = mean(D),
            tests = sum(tests),
            pos = sum(pos),
            pct = pos/tests) %>% 
  mutate(Date = as.Date(date))
```

```{r abm_call_fx}
run_abm <- function(...){
  
# Ages of agents  
  pop.ages <- as.numeric(sample(names(pop.props), size = N/pop.red, replace = TRUE, prob = pop.props))
  
# Baseline risk of infection based on age
  # This gives decent approximation of more risk among age groups that are more active, but could use improvement
  risks <- rgamma(N/pop.red, 2, 2)*(pop.ages+1)^(1/pop.ages)
  
  #hist(risks, breaks = 30, main = "Distribution of individual variability in risk")
  
# Quarantine probabilities 
  q.probs <- rbeta(N/pop.red, 5, 1)
  
  #hist(q.probs, breaks = 30, main = "Distribution of individual probability of quarantining given symptoms")

# Infection status through time
  inf.mat <- matrix(data = NA, nrow = N/pop.red, ncol = t.tot/dt)

# Test status through time
  test.mat <- matrix(data = NA, nrow = N/pop.red, ncol = t.tot/dt)
  
# Event times
  t.til.nxt <- matrix(data = NA, nrow = N/pop.red, ncol = t.tot/dt)
  

# Initial conditions
  e.seed <- 5     #Exposed
  ip.seed <- 1    #infected pre-symptomatic
  ia.seed <- 0    #infected asymptomatic
  im.seed <- 0    #infected mildly symptomatic
  imh.seed <- 0   #infected mildly symptomatic, will become severe
  ih.seed <- 0    #infected severely symptomatic
  d.seed <- 0     #dead
  r.seed <- 0     #removed
  s.seed <- N/pop.red - e.seed - ip.seed - ia.seed - im.seed - imh.seed - ih.seed - d.seed - r.seed

# Initialize infection and waiting time matrices    
init.infection <- sample(c(rep("E", e.seed),
                           rep("Ip", ip.seed),
                           rep("Ia", ia.seed),
                           rep("Im", im.seed),
                           rep("Imh", imh.seed),
                           rep("Ih", ih.seed),
                           rep("D", r.seed),
                           rep("R", r.seed),
                           rep("S", s.seed)), N/pop.red, replace = FALSE)
  
# Fill infection and state transition matrices based on initial conditions  
  inf.mat[,1] <- init.infection  
  t.til.nxt[,1] <- sapply(inf.mat[,1], function(i){
    if(i == "E"){
      t_latent()
    } else if(i == "Ip"){
      t_presymp()
    } else if(i == "Ia"){
      t_asymp()
    } else if(i == "Im"){
      t_msymp()
    } else if(i == "Imh"){
      t_mtosev()
    } else if(i == "Ih"){
      t_sevsymp()
    } else {
      NA
    }
  })
  
for(t in 2:(t.tot/dt)){
  # print(t)
# Advance transition times
    t.til.nxt[,t] <- t.til.nxt[,(t-1)]-dt
  
  # Advance expired states to next state 
    nexts <- which(t.til.nxt[,t] < 0) # states that expired
    if(length(nexts >0)){
      next.mat <- t(mapply(next_state, inf.mat[nexts,(t-1)], pop.ages[nexts])) # function returns new states and time spent in new state
      inf.mat[nexts,t] <- next.mat[,1] # update infection status matrix
      t.til.nxt[nexts,t] <- as.numeric(next.mat[,2]) # update waiting time matrix
    }
    
  # Anyone who hasn't changed infection state remains the same
    sames <- which(is.na(inf.mat[,t]))
    inf.mat[sames,t] <- inf.mat[sames,(t-1)]

  # Implement testing
    n_tests <- round(test_fx(t*dt))
    test.mat[,t] <- test_folks(inf.mat[,t], test.mat[,(t-1)], prob_test, n_tests)
    
  # Implement quarantining  
    quars <- quarantine(inf.mat[,t], test.mat[,t], q.probs)
    
  # Simulate infection
    beta_t <- abm_beta_fx(t*dt)  
    new.Es <- sim.infection(inf.mat[,t], quars, risks, beta_t)

  # Update infection status matrix with new Es  
    inf.mat[new.Es,t] <- "E"
    
  # Update transition times matrix for new Es from sample of latent period dist'n
    t.til.nxt[new.Es,t] <- as.numeric(replicate(length(new.Es), t_latent()))
  
  # On to the next one  
}

pos_tests <- t(apply(test.mat,2,function(x){
  Ttot = sum(x == "T", na.rm = T) ; return(Ttot)
}))[1,]

tests = round(test_fx(1:(t.tot/dt)*dt))

inf_sum <- as.data.frame(sum_inf_mat(inf.mat)) %>% 
  mutate(t = (t.sim*dt),
         date = as.character(t0+t),
         tests = tests,
         tot_pos = pos_tests,
         pos = tot_pos-lag(tot_pos)) %>% 
  group_by(date) %>% 
  summarise(S = mean(S),
            E = mean(E),
            Ip = mean(Ip),
            Ia =mean(Ia),
            Im = mean(Im),
            Imh = mean(Imh),
            Ih = mean(Ih),
            R = mean(R),
            D = mean(D),
            tests = sum(tests),
            pos = sum(pos),
            pct = pos/tests) %>% 
  mutate(Date = as.Date(date))

return(inf_sum)
}
```

```{r abm_replicate, eval = FALSE}
require(parallel)

n.cores <- detectCores()

n.reps <- 15*n.cores

abmclust <- makeCluster(n.cores)         # Make cluster
clusterExport(abmclust, c("t.tot", "t0", "dt", "N", "pop.red",
                          "test_fx", "abm_beta_fx", "n.reps",
                          "pop.props", "prob_test",
                          "sim.infection",
                          "run_abm"))         # Export objects to cluster

clusterEvalQ(abmclust, {
  library(LEMMAABM)
})

abm_sims <- parLapplyLB(abmclust, 1:n.reps, run_abm)

stopCluster(abmclust)

sims <- rep(1:n.reps, each = (t.tot+1))

abm_sims_bind <- bind_rows(abm_sims) %>% 
  mutate(sim = sims)
  

saveRDS(abm_sims_bind, "abm_testing_runs1.rds")
```

### Validate ABM dynamics  
ABM runs in a population an order of magnitude smaller than SF for computational efficiency, so all observed SF data divided by 10 to compare to ABM outputs  

```{r plot_abm_sims_Ih}
abm_sims_bind1 <- readRDS("abm_testing_runs1.rds")
abm_sims_sum1 <- abm_sims_bind1 %>% 
  group_by(date) %>% 
  summarise(S_med = median(S),
            S_25th = quantile(S, 0.25),
            S_75th = quantile(S, 0.75),
            E_med = median(E),
            E_25th = quantile(E, 0.25),
            E_75th = quantile(E, 0.75),
            Ip_med = median(Ip),
            Ip_25th = quantile(Ip, 0.25),
            Ip_75th = quantile(Ip, 0.75),
            Ia_med = median(Ia),
            Ia_25th = quantile(Ia, 0.25),
            Ia_75th = quantile(Ia, 0.75),
            Im_med = median(Im),
            Im_25th = quantile(Im, 0.25),
            Im_75th = quantile(Im, 0.75),
            Imh_med = median(Imh),
            Imh_25th = quantile(Imh, 0.25),
            Imh_75th = quantile(Imh, 0.75),
            Ih_med = median(Ih),
            Ih_25th = quantile(Ih, 0.25),
            Ih_75th = quantile(Ih, 0.75),
            D_med = median(D),
            D_25th = quantile(D, 0.25),
            D_75th = quantile(D, 0.75),
            R_med = median(R, na.rm = T),
            R_25th = quantile(R, 0.25),
            R_75th = quantile(R, 0.75),
            pct_med = median(pct),
            pct_25th = quantile(pct,0.25, na.rm = T),
            pct_75th = quantile(pct, 0.75, na.rm = T)) %>% 
  mutate(Date = as.Date(date))

ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_col(data = sf_all,
             aes(x = Date, y = (HOSP_CONF+ICU_CONF)/pop.red),
           fill = "grey80", col = "black") +
  geom_line(data = abm_sims_sum1,
            aes(x = Date, y = Ih_med),
            col = "red", alpha = 0.5, size = 1.2) +
  geom_ribbon(data = abm_sims_sum1,
              aes(x = Date, ymin = Ih_25th, ymax = Ih_75th),
              fill = "red", alpha = 0.2) +
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
  labs(x = "Date", y = "Hospitalizations/10",
       title = "ABM Model runs comparison to observed data")

```

```{r plot_abm_sims_testing}
ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_col(data = sf_all,
             aes(x = Date, y = pct),
           fill = "grey80", col = "black") +
  geom_line(data = abm_sims_sum1,
            aes(x = Date, y = pct_med),
            col = "blue", alpha = 0.5 , size = 1.2) +
  geom_ribbon(data = abm_sims_sum1,
              aes(x = Date, ymin = pct_25th, ymax = pct_75th),
              fill = "blue", alpha = 0.2) +
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
  labs(x = "Date", y = "% + Tests",
       title = "ABM Testing compared to observed")

```

```{r plot_abm_sims_infections}
abm_sims_Isum1 <- abm_sims_bind1 %>% 
  mutate(I_symp = Im + Imh + Ih,
         I_tot = Ip + Ia + Im + Imh + Ih) %>% 
  group_by(date) %>% 
  summarise(I_symp_med = median(I_symp),
            I_symp_25th = quantile(I_symp, 0.25),
            I_symp_75th = quantile(I_symp, 0.75),
            I_tot_med = median(I_tot),
            I_tot_25th = quantile(I_tot, 0.25),
            I_tot_75th = quantile(I_tot, 0.75),
            pct_med = median(pct),
            pct_25th = quantile(pct,0.25, na.rm = T),
            pct_75th = quantile(pct, 0.75, na.rm = T)) %>% 
  mutate(Date = as.Date(date))

abm_sims_Isum1 %>% 
ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_line(aes(x = Date, y = I_symp_med),
            col = "darkred", alpha = 0.7, size = 1.2) +
  geom_ribbon(aes(x = Date, ymin = I_symp_25th, ymax = I_symp_75th),
              fill = "darkred", alpha = 0.35) +
  geom_line(aes(x = Date, y = I_tot_med),
            col = "orange", alpha = 0.7, size = 1.2) +
  geom_ribbon(aes(x = Date, ymin = I_tot_25th, ymax = I_tot_75th),
              fill = "orange", alpha = 0.35) +
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
  labs(title = "Total and symptomatic infections through time",
       y = "Infections")

```

