---
title: "SiP_Triggers2"
author: "Chris Hoover"
date: "6/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      fig.height = 6,
                      fig.width = 8)

require(deSolve)
require(tidyverse)

```

# Purpose  
Evaluate potential triggers to intervene on COVID transmission (e.g. shelter in place or other method to reduce contact and transmission) from signals in testing data. 

```{r epi_data}
# Case and hosptialization data for SF
sf_hosp <- read.csv(url("https://data.sfgov.org/resource/nxjg-bhem.csv")) %>% 
  mutate(Date = as.Date(reportdate),
         type = ifelse(dphcategory == "ICU", "ICU", "HOSP"),
         conf = ifelse(covidstatus == "PUI", "PUI", "CONF"),
         hosp_stat = paste(type, conf, sep = "_")) %>% 
  pivot_wider(names_from = hosp_stat,
              values_from = patientcount) %>%
  group_by(Date) %>% 
  summarise(ICU_PUI = sum(ICU_PUI, na.rm = T),
            ICU_CONF = sum(ICU_CONF, na.rm = T),
            HOSP_PUI = sum(HOSP_PUI, na.rm = T),
            HOSP_CONF = sum(HOSP_CONF, na.rm = T)) %>% 
  arrange(Date) %>% 
  mutate(HOSP_tot = ICU_CONF + HOSP_CONF,
         HOSP_max = ICU_CONF + HOSP_CONF + ICU_PUI + HOSP_PUI,
         cumICUconf = cumsum(ICU_CONF),
         cumICUpui = cumsum(ICU_PUI),
         cumHOSPconf = cumsum(HOSP_CONF),
         cumHOSPpui = cumsum(HOSP_PUI))

sf_case <- read.csv(url("https://data.sfgov.org/resource/tvq9-ec9w.csv")) %>% 
  mutate(Date = as.Date(date)) %>% 
  pivot_wider(names_from = case_disposition,
              values_from = case_count) %>% 
  group_by(Date) %>% 
  summarise(Cases = sum(Confirmed, na.rm = T),
            Deaths = sum(Death, na.rm = T)) %>% 
  arrange(Date) %>% 
  mutate(cum_case = cumsum(Cases),
         cum_death = cumsum(Deaths))

sf_test <- read.csv(url("https://data.sfgov.org/resource/nfpa-mg4g.csv")) %>% 
  mutate(Date = as.Date(result_date)) %>% 
  arrange(Date) %>% 
  mutate(cum_tests = cumsum(tests),
         cum_pos = cumsum(pos))

sf_all <- sf_test %>% 
  dplyr::select(Date, tests, pos, neg, pct, indeterminate, cum_tests, cum_pos) %>%
  full_join(sf_case, by = "Date") %>% 
  full_join(sf_hosp, by = "Date")
```

## Testing  
```{r test_plot}
sf_test_plot_df <- sf_test %>% 
  mutate(tests_10k = tests/10000,
         tests_7dayavg = zoo::rollmean(tests,7,na.pad = T,align = "right"),
         tests100s_7dayavg = zoo::rollmean(tests/100,7,na.pad = T,align = "right"),
         pos_7day_avg = zoo::rollmean(pos,7,na.pad = T,align = "right")) %>%
  dplyr::select(Date, tests100s_7dayavg, pos_7day_avg) %>% 
  pivot_longer(-Date, names_to = "Test", values_to = "Num")

test_plot <- sf_test_plot_df %>% 
  ggplot() +
    geom_line(aes(x = Date, y = Num, col = Test),
              size = 1.1) +
    theme_bw() +
    scale_color_manual(values = c("purple", "gold"),
                       labels = c("Positive tests",
                                  "Tests\nconducted (100s)")) +
    scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1)) +
    labs(x = "Date",
         y = "Tests and Positives (7-day average)",
         col = "",
         title = "COVID19 testing in SF")

test_plot

```

# Erlang Model  
This is essentially LEMMA with two tweaks: 1) the $E$, $I_R$, $I_H$, and $H$ compartments are erlang distributed to introduce memory in the system in order to account for variability in testing probability, testing sensitivity, etc. and relax the assumption of exponentially distribtued wait times and 2) there is a testing course of infection that mirrors the regular course of infection in order to draw out people who are tested from contributing to infection and to estimate observed vs true measures of the outbreak such as prevalence, attack rate, incidence, case fatality ratio, etc.

\begin{align*}
\dot{S}&=-\beta S\bigg(\sum_{j=1}^{k^{I_R}}I_{R_j}+\sum_{m=1}^{k^{I_H}}I_{H_m}\bigg)/N \\
\dot{E_1}&=\beta S\bigg(\sum_{j=1}^{k^{I_R}}I_{R_j}+\sum_{m=1}^{k^{I_H}}I_{H_m}\bigg)/N-(\sigma+\tau_{E_1})E_1 \\
\dot{E_i}&=\sigma E_{i-1}-(\sigma+\tau_{E_i})E_i\quad,\quad i\in 2,...,k^E \\
\dot{E_1^T}&=\tau_{E_1}E_1-\sigma E_1^T \\
\dot{E_i^T}&=\sigma E_{i-1}^T+\tau_{E_i}E_i-\sigma E_i^T\quad,\quad i\in 2,...,k^E\\
\\
\dot{I_{R_1}}&=\sigma E_{k^E}(1-\alpha)-(\gamma_R+\tau_{I_{R_1}})I_{R_1} \\
\dot{I_{R_j}}&=\gamma_RI_{R_{j-1}}-(\gamma_R+\tau_{I_{R_j}})I_{R_j}\quad,\quad j\in 2,...,k^{I_R} \\
\dot{I_{R_1}^T}&=\tau_{I_{R_1}}I_{R_1}-\gamma_R I_{R_1}^T\\
\dot{I_{R_j}^T}&=\gamma_R I_{R_{j-1}}^T+\tau_{I_{R_j}}I_{R_j}-\gamma_R I_{R_j}^T\quad,\quad j\in 2,...,k^{I_R}\\
\\
\dot{I_{H_1}}&=\sigma E_{k^E}\alpha-(\rho+\tau_{I_{H_1}})I_{H_1} \\
\dot{I_{H_m}}&=\rho I_{H_{m-1}}-(\rho+\tau_{I_{H_m}})I_{H_m}\quad,\quad m\in 2,...,k^{I_H} \\
\dot{I_{H_1}^T}&=\tau_{I_{H_1}}I_{H_1}-\rho I_{H_1}^T\\
\dot{I_{H_m}^T}&=\rho I_{H_{m-1}}^T+\tau_{I_{H_m}}I_{H_m}-\rho I_{H_m}^T\quad,\quad m\in 2,...,k^{I_H}\\
\\
\dot{H^T_1}&=\rho I_{H_{k^{I_H}}}+\rho I_{H_{k^{I_H}}}^T-\gamma_H H_1\\
\dot{H^T_n}&=\gamma_H H_{n-1}-\gamma_H H_n\quad,\quad n\in 2,...,k^H\\
\\
\dot{D}&=\gamma_H\mu H^T_{k^H} \\
\dot{R}&=\gamma_R I_{R_{k^{I_R}}}\\
\dot{R^T}&=\gamma_R I_{R_{k^{I_R}}}^T+\gamma_H(1-\mu)H^T_{k^H}\\
\end{align*}

```{r mod_prep, include=FALSE}
# Key Dates
t0 <- as.Date("2020-02-15")
today <- Sys.Date()

# Beta change dates
int1 <- as.Date("2020-03-05")    # Schools closed
int2 <- as.Date("2020-03-15")    # SiP announced
int3 <- as.Date("2020-04-15")    # Another knot
int4 <- as.Date("2020-05-15")    # Another knot
int5 <- as.Date(Sys.Date()-14)   # Last know two weeks before tooday to reflect hospitalizations as ~2 week lagging indicator
t.end <- as.Date("2020-07-15")   # Where we're headed (for future use)

t.tot <- as.numeric(t.end - t0)

```

```{r mod_pars, include=FALSE}
#Starting covid parameters
pop.size <- 883305
t.sim <- as.numeric(today - t0)

erlang_pars <- list(N = pop.size,     # Total population size
                    t.sim = t.sim,    # Time to run simulation
                    cr = 0.6,         # Relative contact/infectiousness of mild infections
                    ch = 1,           # Relative contact/infectiousness of severe infections
                    E.start = 2,      # Number of exposed at beginning
                    kE = 9,           # Shape (number compartments) for Exposed
                    sigma = 1.5,      # Rate between compartments for Exposed
                    kIr = 3,          # Shape (number compartments) for mild infections
                    gam_r = 0.5,      # Rate between compartments for INfected, will recover
                    kIh = 7,          # Shape (number compartments) for severe infections 
                    rho = 2,          # rate between compartments for Infected, will be hospitalzied
                    kH = 3,           # Shape (number compartments) for Hospitalized
                    gam_h = 0.25,     # Rate between compartments for hospitalized
                    alpha = 0.03,     # Proportion of cases that will end up hospitalized
                    mu = 0.1,         # Proportion of hospitalized cases who will die
                    tau_E = 0,        # Testing rate for exposed
                    tau_Ir = 0,       # Testing rate for mild infections
                    tau_Ih = 0)       # Testing rate for sever infections   

# Transmission parameter estimates
  get_R <- function(beta, pars, S, N){
    with(as.list(pars),{
      ((beta*(1-alpha)*kIr)/gam_r + (beta*alpha*kIh)/rho)*S/N
    })
  }

  get_beta <- function(R, pars){
    with(as.list(pars),{
      (R*gam_r*rho)/((1-alpha)*kIr*rho+alpha*kIh*gam_r)
    })
  }


beta.init <- get_beta(3, erlang_pars)

beta.time <- cbind(-as.numeric(t0-c(t0, int1, int2, int3, int4, int5, t.end)),
                   log(beta.init*c(1, 1, 0.6, 0.4, 0.5, 0.5, 0.7)))

beta_smooth <- splinefun(x = beta.time[,1],
                         y = beta.time[,2],
                         method = "natural")

beta_fx <- approxfun(c(1:t.tot),
                     c(rep(log(beta.init),beta.time[2,1]), sapply(c(beta.time[2,1]+1):t.tot,
                                                                  beta_smooth)))

beta_smooth2 <- approxfun(x = beta.time[,1],
                          y = beta.time[,2],
                          method = "linear")

plot(c(1:150), exp(sapply(1:150, beta_fx)), type = "l")
  lines(c(1:150), exp(sapply(1:150, beta_smooth2)), col = 4)
  points(x = beta.time[,1],
         y = exp(beta.time[,2]),
         pch = 17, col = "red")

```

```{r mod_odes}
erlang_init <- c(pop.size,                       # Starting susceptibles
                 erlang_pars[["E.start"]],       # Starting Exposed, untested class 1
                 rep(0, erlang_pars[["kE"]]-1),  # Starting Exposed, untested classes 2-kE
                 rep(0, erlang_pars[["kIr"]]),   # Starting Ir, untested classes 1-kIr
                 rep(0, erlang_pars[["kIh"]]),   # Starting Ih, untested classes 1-kIh
                 rep(0, erlang_pars[["kE"]]),    # Starting Exposed, tested classes 1-kE
                 rep(0, erlang_pars[["kIr"]]),   # Starting Ir, tested classes 1-kIr
                 rep(0, erlang_pars[["kIh"]]),   # Starting Ih, tested classes 2-kIh
                 rep(0, erlang_pars[["kH"]]),    # Starting Hospitalized, tested classes 1-kH
                 0,0,0)                          # Starting R, Rt and D   

names(erlang_init) <- c("S",                      
                        "E1",       
                        paste0("E", c(2:erlang_pars[["kE"]])),  
                        paste0("Ir", c(1:erlang_pars[["kIr"]])),   # Starting Ir, untested classes 1-kIr
                        paste0("Ih", c(1:erlang_pars[["kIh"]])),   # Starting Ih, untested classes 1-kIh
                        paste0("ET", c(1:(erlang_pars[["kE"]]))),  # Starting Exposed, tested classes 1-kE-1
                        paste0("IrT", c(1:(erlang_pars[["kIr"]]))), # Starting Ir, tested classes 1-kIr-1
                        paste0("IhT", c(1:(erlang_pars[["kIh"]]))), # Starting Ih, tested classes 2-kIh-1
                        paste0("HT", c(1:erlang_pars[["kH"]])),    # Starting Hospitalized, tested classes 1-kH
                        "R","RT","D")                          # Starting R, Rt and D   

infectors <- c(which(names(erlang_init) == "Ir1"):which(names(erlang_init) == paste0("Ih", erlang_pars[["kIh"]])))
rel_inf_rates <- c(rep(erlang_pars[["cr"]], erlang_pars[["kIr"]]),
                   rep(erlang_pars[["ch"]], erlang_pars[["kIh"]]))

erlang_odes = function(t, n, p) { 
  
  # Parameters
    N = p[['N']]             #population size
    kE = p[['kE']]           #Erlang shape for exposed class
    kIr = p[['kIr']]         #Erlang shape for Ir class
    kIh = p[['kIh']]         #Erlang shape for Ih class
    kH = p[['kH']]          #Erlang shape for Th class
    cr = p[['cr']]           #Relative contact rate between S and Ir
    ch = p[['ch']]           #Relative contact rate between S and Ih
    sigma = p[['sigma']]     #erlang scale for E class
    rho = p[['rho']]         #erlang scale for Ih class
    gam_r = p[['gam_r']]     #erlang scale for Ir class
    gam_h = p[['gam_h']]     #erlang scale for H class
    alpha = p[['alpha']]     #proportion severely symptomatic
    mu = p[['mu']]           #proportion of hospitalized cases who die
    tau_E = p[["tau_E"]]     # Testing rate for exposed
    tau_Ir = p[["tau_Ir"]]   # Testing rate for mild infections
    tau_Ih = p[["tau_Ih"]]   # Testing rate for sever infections   


  # Pre-allocate vectors to hold transmission rates for each class
    dS = numeric(1)
    dE = numeric(kE)
    dIr = numeric(kIr)
    dIh = numeric(kIh)
    dR = numeric(1)
    dD = numeric(1)

    dET = numeric(kE)
    dIrT = numeric(kIr)
    dIhT = numeric(kIh)
    dHT = numeric(kH)
    dRT = numeric(1)

  # Beta forcing function for changes in beta over time
    beta = exp(beta_fx(t))
    foi = sum(as.numeric(n[infectors]*rel_inf_rates*beta))/N

  # Transition rates
    #Susceptibles
    dS[1] = -n["S"]*foi
    print(dS)
    #Exposed
    dE[1] = n["S"]*foi - (sigma+tau_E)*n["E1"]
    dE[2:kE] = sapply(2:kE, function(i){
      sigma*n[paste0("E",(i-1))] - (sigma+tau_E)*n[paste0("E",i)]
    })
    
    #Exposed, tested
    dET[1] = tau_E*n["E1"] - sigma*n["ET1"]
    dET[2:kE] = sapply(2:kE, function(i){
      sigma*n[paste0("ET",(i-1))] + tau_E*n[paste0("E",i)] - sigma*n[paste0("ET",i)]
    })
    
    #Infected, mild
    dIr[1] = sigma*n[paste0("E",kE)]*(1-alpha) - (gam_r+tau_Ir)*n["Ir1"]
    dIr[2:kIr] = sapply(2:kIr, function(j){
      gam_r*n[paste0("Ir",(j-1))] - (gam_r+tau_Ir)*n[paste0("Ir",j)]
    })

    #Infected, mild, tested
    dIrT[1] = sigma*n[paste0("ET",kE)]*(1-alpha) + tau_Ir*n["Ir1"] - gam_r*n["Ir1"]
    dIrT[2:kIr] = sapply(2:kIr, function(j){
      gam_r*n[paste0("IrT",(j-1))] + tau_Ir*n[paste0("Ir",j)] - gam_r*n[paste0("IrT",j)]
    })
    
    #Infected, severe
    dIh[1] = sigma*n[paste0("E",kE)]*alpha - (rho+tau_Ih)*n["Ih1"]
    dIh[2:kIh] = sapply(2:kIh, function(l){
      rho*n[paste0("Ih",(l-1))] - (rho+tau_Ih)*n[paste0("Ih",l)]
    })

    #Infected, severe, tested
    dIhT[1] = sigma*n[paste0("ET",kE)]*alpha + tau_Ih*n["Ih1"] - rho*n["Ih1"]
    dIhT[2:kIh] = sapply(2:kIh, function(l){
      rho*n[paste0("IhT",(l-1))] + tau_Ih*n[paste0("Ih",l)] - rho*n[paste0("IhT",l)]
    })
    
    # Hospitalized, tested (Assume everyone is tested once they're admitted to the hospital so no hospitalzed, untested compartment)
    dHT[1] = rho*(n[paste0("Ih",kIh)]+n[paste0("IhT",kIh)]) - gam_h*n["HT1"]
    dHT[2:kH] = sapply(2:kH, function(m){
      gam_h*n[paste0("HT", (m-1))] - gam_h*n[paste0("HT", m)]
    })
    
    # Recovered, tested
    dRT = gam_r*n[paste0("IrT",kIr)] + gam_h*(1-mu)*n[paste0("IhT",kIh)]
    
    # Recovered, never tested
    dR = gam_r*n[paste0("Ir",kIr)] 
    
    # Dead (assume all COVID mortalities are tested)
    dD = gam_h*mu*n[paste0("IhT",kIh)]
    
    out <- c(dS, dE, dIr, dIh, dET, dIrT, dIhT, dHT, dR, dRT, dD)
    out[which(out < 0)] <- 0
    names(out) <- names(n)
    
    #print(out)
    
    return(list(out))
    
} 

erlang_sim <- function(initial_states,
                       sim_time,
                       parameters,
                       ode_fx){
  
  as.data.frame(ode(y = initial_states, 
                    times = 1:sim_time, 
                    parms = parameters,
                    func = ode_fx,
                    method = "impAdams"))
  
}

erlang_test <- erlang_sim(erlang_init,
                          5,
                          erlang_pars,
                          erlang_odes)

# resulting in negative values among IrT and IhT classes
erlang_init2 <- c(pop.size,                       # Starting susceptibles
                  erlang_pars[["E.start"]],       # Starting Exposed, untested class 1
                  rep(1e-3, erlang_pars[["kE"]]-1),  # Starting Exposed, untested classes 2-kE
                  rep(1e-3, erlang_pars[["kIr"]]),   # Starting Ir, untested classes 1-kIr
                  rep(1e-3, erlang_pars[["kIh"]]),   # Starting Ih, untested classes 1-kIh
                  rep(1e-3, erlang_pars[["kE"]]),    # Starting Exposed, tested classes 1-kE
                  rep(1e-3, erlang_pars[["kIr"]]),   # Starting Ir, tested classes 1-kIr
                  rep(1e-3, erlang_pars[["kIh"]]),   # Starting Ih, tested classes 2-kIh
                  rep(1e-3, erlang_pars[["kH"]]),    # Starting Hospitalized, tested classes 1-kH
                  0,0,0)                          # Starting R, Rt and D   

names(erlang_init2) <- c("S",                      
                        "E1",       
                        paste0("E", c(2:erlang_pars[["kE"]])),  
                        paste0("Ir", c(1:erlang_pars[["kIr"]])),   # Starting Ir, untested classes 1-kIr
                        paste0("Ih", c(1:erlang_pars[["kIh"]])),   # Starting Ih, untested classes 1-kIh
                        paste0("ET", c(1:(erlang_pars[["kE"]]))),  # Starting Exposed, tested classes 1-kE-1
                        paste0("IrT", c(1:(erlang_pars[["kIr"]]))), # Starting Ir, tested classes 1-kIr-1
                        paste0("IhT", c(1:(erlang_pars[["kIh"]]))), # Starting Ih, tested classes 2-kIh-1
                        paste0("HT", c(1:erlang_pars[["kH"]])),    # Starting Hospitalized, tested classes 1-kH
                        "R","RT","D")                          # Starting R, Rt and D   


erlang_test2 <- erlang_sim(erlang_init2,
                           5,
                           erlang_pars,
                           erlang_odes)

```

Think the transitions in the above are a bit wonky, so trying to take out simultaneous transitions out of E and into Ir, Ih, Irt, and Iht.

```{r mod_odes2}
erlang_init <- c(pop.size-erlang_pars[["E.start"]], # Starting susceptibles
                 erlang_pars[["E.start"]],          # Starting Exposed, untested class 1
                 rep(0, erlang_pars[["kE"]]-1),    # Starting Exposed, untested classes 2-kE
                 rep(0, erlang_pars[["kIr"]]),     # Starting Ir, untested classes 1-kIr
                 rep(0, erlang_pars[["kIh"]]),     # Starting Ih, untested classes 1-kIh
                 rep(0, erlang_pars[["kE"]]-1),    # Starting Exposed, tested classes 1-kE
                 rep(0, erlang_pars[["kIr"]]-1),   # Starting Ir, tested classes 1-kIr
                 rep(0, erlang_pars[["kIh"]]-1),   # Starting Ih, tested classes 2-kIh
                 rep(0, erlang_pars[["kH"]]),      # Starting Hospitalized, tested classes 1-kH
                 0,0,0)                            # Starting R, Rt and D   

names(erlang_init) <- c("S",                      
                        paste0("E", c(1:erlang_pars[["kE"]])),  
                        paste0("Ir", c(1:erlang_pars[["kIr"]])),   # Starting Ir, untested classes 1-kIr
                        paste0("Ih", c(1:erlang_pars[["kIh"]])),   # Starting Ih, untested classes 1-kIh
                        paste0("ET", c(1:(erlang_pars[["kE"]]-1))),  # Starting Exposed, tested classes 1-kE-1
                        paste0("IrT", c(1:(erlang_pars[["kIr"]]-1))), # Starting Ir, tested classes 1-kIr-1
                        paste0("IhT", c(1:(erlang_pars[["kIh"]]-1))), # Starting Ih, tested classes 2-kIh-1
                        paste0("HT", c(1:erlang_pars[["kH"]])),    # Starting Hospitalized, tested classes 1-kH
                        "R","RT","D")                          # Starting R, Rt and D   

infectors <- c(which(names(erlang_init) == "Ir1"):which(names(erlang_init) == paste0("Ih", erlang_pars[["kIh"]])))
rel_inf_rates <- c(rep(erlang_pars[["cr"]], erlang_pars[["kIr"]]),
                   rep(erlang_pars[["ch"]], erlang_pars[["kIh"]]))

erlang_odes = function(t, n, p) { 
  
  # Parameters
    N = p[['N']]             #population size
    kE = p[['kE']]           #Erlang shape for exposed class
    kIr = p[['kIr']]         #Erlang shape for Ir class
    kIh = p[['kIh']]         #Erlang shape for Ih class
    kH = p[['kH']]          #Erlang shape for Th class
    cr = p[['cr']]           #Relative contact rate between S and Ir
    ch = p[['ch']]           #Relative contact rate between S and Ih
    sigma = p[['sigma']]     #erlang scale for E class
    rho = p[['rho']]         #erlang scale for Ih class
    gam_r = p[['gam_r']]     #erlang scale for Ir class
    gam_h = p[['gam_h']]     #erlang scale for H class
    alpha = p[['alpha']]     #proportion severely symptomatic
    mu = p[['mu']]           #proportion of hospitalized cases who die
    tau_E = p[["tau_E"]]     # Testing rate for exposed
    tau_Ir = p[["tau_Ir"]]   # Testing rate for mild infections
    tau_Ih = p[["tau_Ih"]]   # Testing rate for sever infections   


  # Pre-allocate vectors to hold transmission rates for each class
    dS = numeric(1)
    dE = numeric(kE)
    dIr = numeric(kIr)
    dIh = numeric(kIh)
    dR = numeric(1)
    dD = numeric(1)

    dET = numeric(kE-1)
    dIrT = numeric(kIr-1)
    dIhT = numeric(kIh-1)
    dHT = numeric(kH)
    dRT = numeric(1)

  # Beta forcing function for changes in beta over time
    beta = exp(beta_fx(t))
    foi = sum(as.numeric(n[infectors]*rel_inf_rates*beta))/N

  # Transition rates
    #Susceptibles
    dS[1] = -n["S"]*foi
    
    #Exposed
    dE[1] = n["S"]*foi - (sigma+tau_E)*n["E1"]
    dE[2:(kE-1)] = sapply(2:(kE-1), function(i){
      sigma*n[paste0("E",(i-1))] - (sigma+tau_E)*n[paste0("E",i)]
    })
    dE[kE] = sigma*n[paste0("E",(kE-1))] - sigma*n[paste0("E",kE)]
    
    #Exposed, tested
    dET[1] = tau_E*n["E1"] - sigma*n["ET1"]
    dET[2:(kE-1)] = sapply(2:(kE-1), function(i){
      sigma*n[paste0("ET",(i-1))] + tau_E*n[paste0("E",i)] - sigma*n[paste0("ET",i)]
    })
    
    #Infected, mild
    dIr[1] = sigma*n[paste0("E",kE)]*(1-alpha) - (gam_r+tau_Ir)*n["Ir1"]
    dIr[2:(kIr-1)] = sapply(2:(kIr-1), function(j){
      gam_r*n[paste0("Ir",(j-1))] - (gam_r+tau_Ir)*n[paste0("Ir",j)]
    })
    dIr[kIr] = gam_r*n[paste0("Ir",(kIr-1))] - gam_r*n[paste0("Ir",kIr)]

    #Infected, mild, tested
    dIrT[1] = sigma*n[paste0("ET",kE)]*(1-alpha) + tau_Ir*n["Ir1"] - gam_r*n["Ir1"]
    dIrT[2:(kIr-1)] = sapply(2:(kIr-1), function(j){
      gam_r*n[paste0("IrT",(j-1))] + tau_Ir*n[paste0("Ir",j)] - gam_r*n[paste0("IrT",j)]
    })
    
    #Infected, severe
    dIh[1] = sigma*n[paste0("E",kE)]*alpha - (rho+tau_Ih)*n["Ih1"]
    dIh[2:(kIh-1)] = sapply(2:(kIh-1), function(l){
      rho*n[paste0("Ih",(l-1))] - (rho+tau_Ih)*n[paste0("Ih",l)]
    })
    dIh[kIh] = rho*n[paste0("Ih",(kIh-1))] - rho*n[paste0("Ih",kIh)]

    #Infected, severe, tested
    dIhT[1] = sigma*n[paste0("ET",kE)]*alpha + tau_Ih*n["Ih1"] - rho*n["Ih1"]
    dIhT[2:(kIh-1)] = sapply(2:(kIh-1), function(l){
      rho*n[paste0("IhT",(l-1))] + tau_Ih*n[paste0("Ih",l)] - rho*n[paste0("IhT",l)]
    })
    
    # Hospitalized, tested (Assume everyone is tested once they're admitted to the hospital so no hospitalzed, untested compartment)
    dHT[1] = rho*(n[paste0("Ih",kIh)]+n[paste0("IhT",kIh)]) - gam_h*n["HT1"]
    dHT[2:kH] = sapply(2:kH, function(m){
      gam_h*n[paste0("HT", (m-1))] - gam_h*n[paste0("HT", m)]
    })
    
    # Recovered, tested
    dRT = gam_r*n[paste0("IrT",kIr)] + gam_h*(1-mu)*n[paste0("IhT",kIh)]
    
    # Recovered, never tested
    dR = gam_r*n[paste0("Ir",kIr)] 
    
    # Dead (assume all COVID mortalities are tested)
    dD = gam_h*mu*n[paste0("IhT",kIh)]
    
    out <- c(dS, dE, dIr, dIh, dET, dIrT, dIhT, dHT, dR, dRT, dD)

    #out[which(out < 0)] <- 0
    names(out) <- names(n)
    
    #print(out)
    
    return(list(out))
    
} 

erlang_sim <- function(initial_states,
                       sim_time,
                       parameters,
                       ode_fx){
  
  as.data.frame(ode(y = initial_states, 
                    times = 1:sim_time, 
                    parms = parameters,
                    func = ode_fx,
                    method = "impAdams"))
  
}

erlang_test <- erlang_sim(erlang_init,
                          5,
                          erlang_pars,
                          erlang_odes)

# resulting in negative values among IrT and IhT classes
erlang_init2 <- c(pop.size - erlang_pars[["E.start"]],# Starting susceptibles
                  erlang_pars[["E.start"]],           # Starting Exposed, untested class 1
                  rep(1e-3, erlang_pars[["kE"]]-1),  # Starting Exposed, untested classes 2-kE
                  rep(1e-3, erlang_pars[["kIr"]]),   # Starting Ir, untested classes 1-kIr
                  rep(1e-3, erlang_pars[["kIh"]]),   # Starting Ih, untested classes 1-kIh
                  rep(1e-3, erlang_pars[["kE"]]),    # Starting Exposed, tested classes 1-kE
                  rep(1e-3, erlang_pars[["kIr"]]),   # Starting Ir, tested classes 1-kIr
                  rep(1e-3, erlang_pars[["kIh"]]),   # Starting Ih, tested classes 2-kIh
                  rep(1e-3, erlang_pars[["kH"]]),    # Starting Hospitalized, tested classes 1-kH
                  0,0,0)                          # Starting R, Rt and D   

names(erlang_init2) <- c("S",                      
                        "E1",       
                        paste0("E", c(2:erlang_pars[["kE"]])),  
                        paste0("Ir", c(1:erlang_pars[["kIr"]])),   # Starting Ir, untested classes 1-kIr
                        paste0("Ih", c(1:erlang_pars[["kIh"]])),   # Starting Ih, untested classes 1-kIh
                        paste0("ET", c(1:(erlang_pars[["kE"]]))),  # Starting Exposed, tested classes 1-kE-1
                        paste0("IrT", c(1:(erlang_pars[["kIr"]]))), # Starting Ir, tested classes 1-kIr-1
                        paste0("IhT", c(1:(erlang_pars[["kIh"]]))), # Starting Ih, tested classes 2-kIh-1
                        paste0("HT", c(1:erlang_pars[["kH"]])),    # Starting Hospitalized, tested classes 1-kH
                        "R","RT","D")                          # Starting R, Rt and D   


erlang_test2 <- erlang_sim(erlang_init2,
                           5,
                           erlang_pars,
                           erlang_odes)

```


```{r mod_test2}
erlang_pars_test <- erlang_pars
erlang_pars_test[["tau_E"]] <- 1600/pop.size*0.1
erlang_pars_test[["tau_Ir"]] <- 1600/pop.size*0.3
erlang_pars_test[["tau_Ih"]] <- 1600/pop.size*0.6

erlang_test_the_testing <- erlang_sim(erlang_init,
                                      t.tot,
                                      erlang_pars_test,
                                      erlang_odes)

```

# Lemma Model with testing  
```{r lemma_test_odes}
lemma_test_pars <- list(N = pop.size,
                   t.sim = t.sim,
                   E.start = 2,
                   cr = 1,
                   ch = 1,
                   sigma = 1/3,
                   alpha = 0.05,
                   rho = 1/4,
                   gam_r = 1/5,
                   gam_h = 1/12,
                   mu = 0.2,
                   tau_E = 0,
                   tau_Ir = 0,
                   tau_Ih = 0)

E.start = 2

lemma_init <- c(S = pop.size - E.start,
                E = E.start,
                Ir = 0,
                Ih = 0,
                ET = 0,
                IrT = 0,
                IhT = 0,
                H = 0,
                D = 0,
                R = 0,
                RT = 0)


lemma_test_odes = function(t, n, p, fx) { 

  # States
    S = n[1]
    E = n[2]
    Ir = n[3]
    Ih = n[4]
    ET = n[5]
    IrT = n[6]
    IhT = n[7]
    H = n[8]
    D = n[9]
    R = n[10]
    RT = n[11]
  
  # Parameters
    N = p[['N']]             #population size
    cr = p[['cr']]           #Relative contact rate between S and Ir
    ch = p[['ch']]           #Relative contact rate between S and Ih
    sigma = p[['sigma']]     #1/serial interval
    alpha = p[['alpha']]     #proportion severely symptomatic
    rho = p[['rho']]         #1/time between symptom onset and hospitalization
    gam_r = p[['gam_r']]     #1/time to recovery (non-infectiousness) for mild/asymptomatic
    gam_h = p[['gam_h']]     #1/time to removed (recovered or deceased) for hospitalized
    mu = p[['mu']]           #proportion of hospitalized cases who die
    tau_E = p[["tau_E"]]     # Testing rate for exposed
    tau_Ir = p[["tau_Ir"]]   # Testing rate for mild infections
    tau_Ih = p[["tau_Ih"]]   # Testing rate for sever infections   
beta = 0.52
  #Movement forcing function
    #beta = exp(fx(t))
  
  # Model equations
    dSdt = -S*beta*(Ir*cr+Ih*ch)/N
    
    dEdt = S*beta*(Ir*cr+Ih*ch)/N - E*(sigma+tau_E)
    
    dIrdt = E*(1-alpha)*sigma - Ir*(gam_r+tau_Ir)
    
    dIhdt = E*alpha*sigma - Ih*(rho*tau_Ih)

    dETdt = tau_E*E - ET*sigma
    
    dIrTdt = ET*(1-alpha)*sigma - IrT*gam_r
    
    dIhTdt = ET*alpha*sigma - IhT*rho

    dHdt = (Ih+IhT)*rho - H*gam_h
    
    dDdt = H*gam_h*mu
    
    dRdt = Ir*gam_r
    
    dRTdt = IrT*gam_r+H*gam_h*(1-mu)

    return(list(c(dSdt, dEdt, dIrdt, dIhdt, dETdt, dIrTdt, dIhTdt, dHdt, dDdt, dRdt, dRTdt)))
    
} 

lemme_test_sim <- as.data.frame(ode(y = lemma_init, 
                                    times = 1:t.tot, 
                                    parms = lemma_test_pars,
                                    func = lemma_test_odes, 
                                    fx = beta_fx))

```