---
title: "COVID Testing Bias and Influence on Projected Testing metrics"
author: "Chris Hoover"
date: "6/4/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 6,
                      fig.width = 8)

require(parallel)
require(deSolve)
require(wrswoR)
require(zoo)
require(tidyverse)
devtools::load_all("../ABM")

```

# Purpose  
Estimate bias in COVID19 testing in SF and incorporate into near-term projections to determine influence on testing metrics such as %+ tests. 

```{r epi_data}
# Case and hosptialization data for SF
sf_hosp <- read.csv(url("https://data.sfgov.org/resource/nxjg-bhem.csv")) %>% 
  mutate(Date = as.Date(reportdate),
         type = ifelse(dphcategory == "ICU", "ICU", "HOSP"),
         conf = ifelse(covidstatus == "PUI", "PUI", "CONF"),
         hosp_stat = paste(type, conf, sep = "_")) %>% 
  pivot_wider(names_from = hosp_stat,
              values_from = patientcount) %>%
  group_by(Date) %>% 
  summarise(ICU_PUI = sum(ICU_PUI, na.rm = T),
            ICU_CONF = sum(ICU_CONF, na.rm = T),
            HOSP_PUI = sum(HOSP_PUI, na.rm = T),
            HOSP_CONF = sum(HOSP_CONF, na.rm = T)) %>% 
  arrange(Date) %>% 
  mutate(HOSP_tot = ICU_CONF + HOSP_CONF,
         HOSP_max = ICU_CONF + HOSP_CONF + ICU_PUI + HOSP_PUI,
         cumICUconf = cumsum(ICU_CONF),
         cumICUpui = cumsum(ICU_PUI),
         cumHOSPconf = cumsum(HOSP_CONF),
         cumHOSPpui = cumsum(HOSP_PUI))

sf_case <- read.csv(url("https://data.sfgov.org/resource/tvq9-ec9w.csv")) %>% 
  mutate(Date = as.Date(specimen_collection_date)) %>% 
  pivot_wider(names_from = case_disposition,
              values_from = case_count) %>% 
  group_by(Date) %>% 
  summarise(Cases = sum(Confirmed, na.rm = T),
            Deaths = sum(Death, na.rm = T)) %>% 
  arrange(Date) %>% 
  mutate(cum_case = cumsum(Cases),
         cum_death = cumsum(Deaths))

sf_test <- read.csv(url("https://data.sfgov.org/resource/nfpa-mg4g.csv")) %>% 
  mutate(Date = as.Date(specimen_collection_date)) %>% 
  arrange(Date) %>% 
  mutate(cum_tests = cumsum(tests),
         cum_pos = cumsum(pos))

sf_all <- sf_test %>% 
  dplyr::select(Date, tests, pos, neg, pct, indeterminate, cum_tests, cum_pos) %>%
  full_join(sf_case, by = "Date") %>% 
  full_join(sf_hosp, by = "Date")
```

## Testing  
```{r test_plot}
sf_test_plot_df <- sf_test %>% 
  mutate(tests_10k = tests/10000,
         tests_7dayavg = zoo::rollmean(tests,7,na.pad = T,align = "right"),
         tests100s_7dayavg = zoo::rollmean(tests/100,7,na.pad = T,align = "right"),
         pos_7day_avg = zoo::rollmean(pos,7,na.pad = T,align = "right")) %>%
  dplyr::select(Date, tests100s_7dayavg, pos_7day_avg) %>% 
  pivot_longer(-Date, names_to = "Test", values_to = "Num")

test_plot <- sf_test_plot_df %>% 
  ggplot() +
    geom_line(aes(x = Date, y = Num, col = Test),
              size = 1.1) +
    theme_bw() +
    scale_color_manual(values = c("purple", "gold"),
                       labels = c("Positive tests",
                                  "Tests\nconducted (100s)")) +
    scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1)) +
    labs(x = "Date",
         y = "Tests and Positives (7-day average)",
         col = "",
         title = "COVID19 testing in SF")

test_plot

```

# Model  
We use a slight tweak to LEMMA to add an explicit deaths compartment in order to fit to deaths data in addition to hospitalizations  

\begin{align*}
\dot{S}&=-\beta S(I_R+I_H)/N \\
\dot{E}&=\beta S(I_R+I_H)/N-\sigma E \\
\dot{I_R}&=\sigma(1-\alpha) E-\gamma_R I_R \\
\dot{I_H}&=\sigma\alpha E-\rho I_H \\
\dot{H}&=\rho I_H-\gamma_H H \\
\dot{D}&=\gamma_H\mu H \\
\dot{R}&=\gamma_R I_R \gamma_H(1-\mu)H\\
\end{align*}

```{r mod_prep, include=FALSE}
# Key Dates
t0 <- as.Date("2020-02-15")
last_dat <- max(sf_all %>% na.omit() %>% pull(Date))
today <- Sys.Date()

# Beta change dates
db1 <- as.Date("2020-03-05") #Schools closed
db2 <- as.Date("2020-03-15")     #SiP announced
db3 <- as.Date("2020-04-15")          #Another knot
db4 <- as.Date("2020-05-15")          #Another knot
last_fit <- last_dat-7
t.end <- as.Date("2020-07-15")         # Where we're headed (for future use)

t.tot <- as.numeric(t.end - t0)

# Transmission parameter estimates
get_R <- function(beta, alpha, gam_r, gam_h, S, N){
  ((beta*(1-alpha))/gam_r + (beta*alpha)/gam_h)*S/N
}

get_beta <- function(R, alpha, gam_r, gam_h){
  (R*gam_r*gam_h)/((1-alpha)*gam_h+alpha*gam_r*gam_h)
}

#Starting covid parameters
pop.size <- 883305
t.sim <- as.numeric(last_fit - t0)

lemma_pars <- list(N = pop.size,
                   t.sim = t.sim,
                   E.start = 2,
                   cr = 1,
                   ch = 1,
                   sigma = 1/3,
                   alpha = 0.05,
                   rho = 1/4,
                   gam_r = 1/5,
                   gam_h = 1/12,
                   mu = 0.2)

beta.init <- get_beta(3, lemma_pars[["alpha"]], lemma_pars[["gam_r"]], lemma_pars[["rho"]])

beta.time <- cbind(-as.numeric(t0-c(t0, db1,
                                    db2, db3, db4, last_fit, t.end)),
                   log(beta.init*c(1, 1, 0.6, 0.4, 0.5, 0.5, 0.7)))

beta_smooth <- splinefun(x = beta.time[,1],
                         y = beta.time[,2],
                         method = "natural")

beta_fx <- approxfun(c(1:t.tot),
                     c(rep(log(beta.init),beta.time[2,1]), sapply(c(beta.time[2,1]+1):t.tot,
                                                                  beta_smooth)))

beta_smooth2 <- approxfun(x = beta.time[,1],
                          y = beta.time[,2],
                          method = "linear")

plot(c(1:t.tot), exp(sapply(1:t.tot, beta_fx)), type = "l")
  lines(c(1:t.tot), exp(sapply(1:t.tot, beta_smooth2)), col = 4)
  points(x = beta.time[,1],
         y = exp(beta.time[,2]),
         pch = 17, col = "red")

```

```{r mod_odes}
lemma_odes = function(t, n, p, fx) { 

  # States
    S = n[1]
    E = n[2]
    Ir = n[3]
    Ih = n[4]
    H = n[5]
    D = n[6]
    R = n[7]
  
  # Parameters
    N = p[['N']]             #population size
    cr = p[['cr']]           #Relative contact rate between S and Ir
    ch = p[['ch']]           #Relative contact rate between S and Ih
    sigma = p[['sigma']]     #1/serial interval
    alpha = p[['alpha']]     #proportion severely symptomatic
    rho = p[['rho']]         #1/time between symptom onset and hospitalization
    gam_r = p[['gam_r']]     #1/time to recovery (non-infectiousness) for mild/asymptomatic
    gam_h = p[['gam_h']]     #1/time to removed (recovered or deceased) for hospitalized
    mu = p[['mu']]           #proportion of hospitalized cases who die

  #Movement forcing function
    beta = exp(fx(t))
  
  # Model equations
    dSdt = -S*beta*(Ir*cr+Ih*ch)/N
    
    dEdt = S*beta*(Ir*cr+Ih*ch)/N - E*sigma
    
    dIrdt = E*(1-alpha)*sigma - Ir*gam_r
    
    dIhdt = E*alpha*sigma - Ih*rho

    dHdt = Ih*rho - H*gam_h
    
    dDdt = H*gam_h*mu
    
    dRdt = Ir*gam_r+H*gam_h*(1-mu)

    return(list(c(dSdt, dEdt, dIrdt, dIhdt, dHdt, dDdt, dRdt)))
    
} 
```

## Model fit  

```{r fit_fx}
#Function to return fit from input parameters and data
  lemma_fit <- function(fit.pars, all.pars, fit_data) {
    use.pars <- all.pars
    use.pars[["alpha"]] <- fit.pars[3]
    use.pars[["mu"]] <- fit.pars[4]
        
    beta_smooth <- splinefun(x = beta.time[2:6,1],
                             y = log(c(fit.pars[2],
                                       fit.pars[5],
                                       fit.pars[6],
                                       fit.pars[7],
                                       fit.pars[8])),
                             method = "natural")
  
    beta_fx <- approxfun(c(1:use.pars[["t.sim"]]),
                         c(log(rep(fit.pars[2],beta.time[2,1])), 
                           sapply(c(beta.time[2,1]+1):use.pars[["t.sim"]],
                                  beta_smooth)))

    # initial number of exposed as parameter
    init <- c(S = as.numeric(use.pars[["N"]] - fit.pars[1]),
              E = as.numeric(fit.pars[1]),
              Ir = 0,
              Ih = 0,
              H = 0,
              D = 0,
              R = 0)
    
    # Run the model with input parameters to generate predicted quantities
    mod.run <- as.data.frame(ode(y = init, 
                                 times = 1:use.pars[["t.sim"]], 
                                 parms = use.pars,
                                 func = lemma_odes, 
                                 fx = beta_fx)) %>% 
      mutate(Date = as.Date(t0+(time-1)))
    
    # Join simulated and observed data
    mod.obs.dat <- mod.run %>% 
      full_join(fit_data, by = "Date")
    
    ## Calculate the log likelihood (logLike) of the data given theta (parameter set)
  	hosp.sse <- (mod.obs.dat$HOSP_CONF+mod.obs.dat$ICU_CONF - mod.obs.dat$H)^2 
  	mort.sse <- (mod.obs.dat$D - mod.obs.dat$cum_death)^2

  	return(sum(hosp.sse+mort.sse, na.rm = T))
  }

#Function to return fit from input log parameters and data
  lemma_fit_log <- function(fit.pars, all.pars, fit_data) {
    use.pars <- all.pars
    use.pars[["alpha"]] <- exp(fit.pars[3])
    use.pars[["mu"]] <- exp(fit.pars[4])
    
    beta_smooth <- splinefun(x = beta.time[2:6,1],
                             y = log(exp(c(fit.pars[2],
                                           fit.pars[5],
                                           fit.pars[6],
                                           fit.pars[7],
                                           fit.pars[8]))),
                             method = "natural")
  
    beta_fx <- approxfun(c(1:use.pars[["t.sim"]]),
                         c(log(rep(exp(fit.pars[2]),beta.time[2,1])), 
                           sapply(c(beta.time[2,1]+1):use.pars[["t.sim"]],
                                  beta_smooth)))

    # initial number of exposed as parameter
    init <- c(S = use.pars[["N"]] - exp(fit.pars[1]),
              E = exp(fit.pars[1]),
              Ir = 0,
              Ih = 0,
              H = 0,
              D = 0,
              R = 0)
    
    # Run the model with input parameters to generate predicted quantities
    mod.run <- as.data.frame(ode(y = init, 
                                 times = 1:use.pars[["t.sim"]], 
                                 parms = use.pars,
                                 func = lemma_odes, 
                                 fx = beta_fx)) %>% 
      mutate(Date = as.Date(t0+(time-1)))
    
    # Join simulated and observed data
    mod.obs.dat <- mod.run %>% 
      full_join(fit_data, by = "Date")
    
    ## Calculate the log likelihood (logLike) of the data given theta (parameter set)
  	hosp.sse <- (mod.obs.dat$HOSP_CONF+mod.obs.dat$ICU_CONF - mod.obs.dat$H)^2 
  	mort.sse <- (mod.obs.dat$D - mod.obs.dat$cum_death)^2

  	return(sum(hosp.sse+mort.sse, na.rm = T))
  }

```

```{r fit, eval = FALSE}
iterations <- 1000

set.seed(430)

E.start = rnbinom(iterations, mu = 3, size = 30)+1
beta = rbeta(iterations, 4, 2)
alpha = rbeta(iterations, 2,40)
mu = rbeta(iterations, 3, 12)
db1 = beta*rbeta(iterations, 4, 2)       # Change in transmission from school closure to SiP
db2 = db1*rbeta(iterations, 4, 2)        # Change in transmission 4-15
db3 = db2*(1-rnorm(iterations, 0, 0.2))  # Change in transmission 5-15
db4 = db3*(1-rnorm(iterations, 0, 0.1))  # Change in transmission at last data point

prop.pars <- cbind(E.start, beta, alpha, mu,
                   db1, db2, db3, db4)

lemma_sweeps <- as.list(data.frame(t(prop.pars)))

start <- Sys.time()
lemma_fits <- lapply(lemma_sweeps, lemma_fit,
                     lemma_pars, sf_all)
end <- Sys.time()

end-start

lemma_rslts <- cbind(prop.pars, t(data.frame(lemma_fits)))

```

```{r fit_parallel, eval = FALSE}
iterations <- 10000

set.seed(430)

E.start = rnbinom(iterations, mu = 3, size = 30)+1
beta = rbeta(iterations, 4, 2)
alpha = runif(iterations, 0.02, 0.04)
mu = runif(iterations, 0.08, 0.14)
db1 = beta*rbeta(iterations, 4, 2)       # Change in transmission from school closure to SiP
db2 = db1*rbeta(iterations, 4, 2)        # Change in transmission 4-15
db3 = db2*(1-rnorm(iterations, 0, 0.2))  # Change in transmission 5-15
db4 = db3*(1-rnorm(iterations, 0, 0.2))  # Change in transmission at last data point

prop.pars <- cbind(E.start, beta, alpha, mu,
                   db1, db2, db3, db4)

as.data.frame(prop.pars) %>% 
  pivot_longer(E.start:db4, names_to = "Par") %>% 
  ggplot() +
    geom_density(aes(value, fill = Par), alpha = 0.5) +
    facet_wrap(Par~., ncol = 4, nrow = 2,
               scales = "free")

lemma_sweeps <- as.list(data.frame(t(prop.pars)))

require(parallel)
opclust <- makeCluster(detectCores())         # Make cluster
clusterExport(opclust, c("lemma_sweeps",
                         "lemma_pars", 
                         "lemma_fit",
                         "sf_all",
                         "lemma_odes",
                         "beta.time",
                         "t0"))         # Export objects to cluster

clusterEvalQ(opclust, {
  library(deSolve)
  library(tidyverse)
})

lemma_fits <- parLapply(opclust, lemma_sweeps, lemma_fit, 
                        lemma_pars, sf_all)

stopCluster(opclust)

lemma_rslts <- cbind(prop.pars, t(data.frame(lemma_fits)))

saveRDS(lemma_rslts, 
        "LEMMA_ODE_sweeps_6-15-20.rds")
```

```{r improve_fit, eval = FALSE}
lemma_rslts <- readRDS("LEMMA_ODE_sweeps_6-15-20.rds")

best_pt <- optim(par = log(lemma_rslts[which(lemma_rslts[,9] == min(lemma_rslts[,9])),][-9]),
                 fn = lemma_fit_log,
                 all.pars = lemma_pars,
                 fit_data = sf_all,
                 method = "L-BFGS-B",
                 lower = log(c(1, 1e-2, 0.02, 0.08,1e-2,1e-2,1e-2,1e-2)),
                 upper = log(c(15, 0.9, 0.04, 0.14,1,1,1,1)),
                 control = list(maxit = 1000,
                                trace = 1))

best_pt_pars <- exp(best_pt$par)

saveRDS(best_pt_pars, "LEMMA_ODE_best_fit_pars_6-15-20.rds")
```

```{r sim_best, include = FALSE}
best_pt_pars <- readRDS("LEMMA_ODE_best_fit_pars_6-15-20.rds")

best.pars <- lemma_pars
best.pars[["E.start"]] <- round(as.numeric(best_pt_pars[1]))
best.pars[["alpha"]] <- as.numeric(best_pt_pars[3])
best.pars[["mu"]] <- as.numeric(best_pt_pars[4])

  init <- c(S = as.numeric(best.pars[["N"]] - round(best.pars[["E.start"]])),
            E = as.numeric(round(best.pars[["E.start"]])),
            Ir = 0,
            Ih = 0,
            H = 0,
            D = 0,
            R = 0)
  
  best_spline <- splinefun(x = beta.time[1:6,1],
                           y = log(c(best_pt_pars[2],best_pt_pars[2],
                                     as.numeric(best_pt_pars[5:8]))),
                           method = "natural")
  
  best_smooth <- approxfun(c(1:best.pars[["t.sim"]]),
                           c(rep(log(best_pt_pars[2]),beta.time[2,1]), 
                             sapply(c(beta.time[2,1]+1):best.pars[["t.sim"]],
                                    best_spline)),
                           method = "constant")

plot(c(1:best.pars[["t.sim"]]), exp(sapply(1:best.pars[["t.sim"]], best_smooth)), 
     type = "l")
  points(x = beta.time[1:6,1],
         y = c(best_pt_pars[2],best_pt_pars[2],
                   as.numeric(best_pt_pars[5:8])),
         pch = 17, col = "red")
```

```{r best_pars, dev="tikz" }
tab1 <- data.frame(val = as.character(round(as.numeric(best.pars), 3)), 
                   des = c("population size",
                           "time to run simulation",
                           "starting number of exposed",
                           "Relative contact rate between S and Ir",
                           "Relative contact rate between S and Ih",
                           "1/serial interval",
                           "proportion severely symptomatic (will be hospitalized)",
                           "time between symptom onset and hospitalization",
                           "1/time to recovery (non-infectiousness) for mildly symptomatic",
                           "1/time hospitalized",
                           "proportion of hospitalized cases who die"))

rownames(tab1) <- c("$N$", "t.sim", "$E_0$" ,"$c_r$","$c_h$",
                    "$\\sigma$","$\\alpha$", "$\\rho$", 
                    "$\\gamma_r$","$\\gamma_h$","$\\mu$")

knitr::kable(tab1, row.names = TRUE, 
             col.names = c("Value", "Definition"), 
             format = "latex", escape = FALSE, 
             caption = "Best fit model parameters")
```

```{r plot_best}
best.sim <- as.data.frame(ode(y = init, 
                              times = 1:best.pars[["t.sim"]], 
                              parms = best.pars,
                              func = lemma_odes, 
                              fx = best_smooth)) %>% 
  mutate(Date = as.Date(t0+time-1),
         I_tot = Ir+Ih) 

best.sim$Re <- get_R(exp(sapply(1:best.pars[["t.sim"]], best_smooth)),
                     best.pars[["alpha"]], best.pars[["gam_r"]], best.pars[["rho"]], best.sim$S, best.pars[["N"]])

ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_col(data = sf_all,
             aes(x = Date, y = HOSP_CONF+ICU_CONF),
           col = "black", fill = "grey50") +
  geom_line(data = best.sim,
            aes(x = Date, y = H),
            col = "coral", size = 1.2) +
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
  labs(x = "Date", y = "Hospitalizations",
       title = "Model fit to SF hospitalizations")

ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_col(data = sf_all,
             aes(x = Date, y = cum_death),
           col = "black", fill = "grey50") +
  geom_line(data = best.sim,
            aes(x = Date, y = D),
            col = "coral", size = 1.2) +
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
  labs(x = "Date", y = "Cumulative deaths",
       title = "Model fit to SF deaths")

ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_line(data = best.sim,
            aes(x = Date, y = Re),
            col = "coral", size = 1.2) +
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
  scale_y_continuous(breaks = seq(0,3.5,0.5),
                     limits = c(0,4)) +
  geom_hline(yintercept = 1, lty = 2) +
  labs(x = "Date", y = expression(R[e]),
       title = "Best fit effective reproduction number")

```

### Comparison to testing data  
```{r comp_test}
sf_test_p_inc <- rbind(sf_test_plot_df,
      best.sim %>% 
        mutate(Test = " Modeled Cases",
               Num = I_tot/best.pars[["N"]]*100000) %>% 
        dplyr::select(Date,Test,Num))

sf_test_p_inc %>% 
  ggplot() +
    geom_line(aes(x = Date, y = Num, col = Test),
              size = 1.1) +
    theme_bw() +
    scale_color_manual(values = c("darkblue", "purple", "gold"),
                       labels = c("Modeled\nCases/100k",
                                  "n positive",
                                  "Tests\n conducted (100s)")) +
    scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1)) +
    labs(x = "Date", y = "",
         title = "SF COVID testing and modeled cases",
         col = "")

```

```{r test_mod_comp}
sf_test_mod_join <- sf_test %>% 
  left_join(best.sim, by = "Date") %>% 
  mutate(tests_7dayavg = zoo::rollmean(tests,7,na.pad = T,align = "right"),
         pos_7day_avg = zoo::rollmean(pos,7,na.pad = T,align = "right"),
         H_new = Ih*best.pars[["rho"]],
         I_new = E*best.pars[["sigma"]],
         Ir_new = E*(1-best.pars[["alpha"]])*best.pars[["sigma"]],
         Ih_new = E*best.pars[["alpha"]]*best.pars[["sigma"]])

sf_test_mod_join %>% 
  mutate(tests_100_avg = tests_7dayavg/100,
         pos_p1000_avg = pos_7day_avg/1000,
         H_new_p100 = H_new/100,
         I_new_p1000 = I_new/1000) %>% 
  dplyr::select(Date, tests_100_avg, pos_7day_avg, H_new, I_new) %>% 
  pivot_longer(-Date, names_to = "Var") %>%
  ggplot() +
    geom_line(aes(x = Date, y = value, col = Var),
              size = 1.1) +
    theme_bw() +
    scale_color_manual(values = c("darkred", "darkgreen", "purple", "gold"),
                       labels = c("Mod H admissions",
                                  "Mod I incidence",
                                  "Positives",
                                  "Tests (100s)")) +
    scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1)) +
    labs(x = "Date", y = "",
         title = "SF COVID testing and model",
         col = "")

```

#### Match modeled testing to observed testing  
Assume all new hospitalized cases are tested and confirmed positive, then assume remaining tests are allocated to non-hospitalized population. In the model, this is equivalent to a sample of $S$, $E$, and $I_r+I_h$. In reality, lots of nuance in the $E$ and $R$ compartments with regard to testing, but for simplicity, we'll assume positive tests from this sample only come from $I_r+I_h$ and $R$s are not tested. So want to solve for sampling bias, $\mathcal{B}$, from:
$$\frac{+Tests-H_{new}}{Tests-H_{new}}=\frac{\mathcal{B}(I_r+I_h)}{S+E+I_r+I_h}$$

where $\mathcal{B}$ can be roughly interpreted as the enhanced probability of infectious individuals being tested

```{r mod_test_match, fig.cap="Distribution of estimated sampling bias since end of April"}
sf_test_mod_join_samp_bias <- sf_test_mod_join %>% 
  mutate(tests_100_avg = tests_7dayavg/100,
         mod_SE = S+E,
         tests_non_H = tests - round(H_new),
         pos_non_H = pos - round(H_new),
         non_H_pos_rate = pos_non_H/tests_non_H,
         mod_I_prev = I_tot/(mod_SE+I_tot),
         bias = non_H_pos_rate/mod_I_prev)

boxplot(sf_test_mod_join_samp_bias %>% 
          filter(Date >= as.Date("2020-04-30")) %>% 
          pull(bias),
        ylab = "Bias")

bias_median <- median(sf_test_mod_join_samp_bias %>% 
                        filter(Date >= as.Date("2020-04-30")) %>% 
                        pull(bias), na.rm = T)

```

```{r bias_through_time}
sf_test_bias <- sf_test_mod_join_samp_bias %>% 
  mutate(bias_7dayavg = zoo::rollmean(bias, 7, fill = NA)) %>% 
  filter(Date >= (as.Date("2020-03-15"))) %>% 
  ggplot() +
    geom_line(aes(x = Date, y = bias_7dayavg),
              size = 1.1) +
    theme_bw() +
    scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day") +
    scale_y_continuous(breaks = c(0, 25,50,75,100,125, 150),
                       limits = c(0,150)) +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1)) +
    labs(x = "Date", y = "Testing bias (7-day average)",
         title = "SF COVID testing bias through time")

sf_test_bias
```


```{r sf_covid_testing_bias_and_numbers}
sf_test <- sf_test %>% 
  mutate(tests_p100k = (tests/pop.size)*1e5,
         tests_p100k_smooth = rollmean(tests_p100k, 7, fill = NA, align = "left"))

sf_test_bias +
  geom_line(data = sf_test,
            aes(x = Date, y = tests_p100k_smooth), 
            col = "purple", size = 1.2) +
    ylim(c(0,300)) +
    labs(x = "Date", y = "Tests/100k and bias (7-day average)",
         title = "SF COVID Testing Numbers and Estimated Bias")
  

```


```{r test_mod_test_match, eval = FALSE}
#Simulate biased testing of modeled data
test_with_bias <- function(p_tests, n_tests, mod_I, mod_H, mod_SE){
  B <- (p_tests-round(mod_H))/(n_tests-round(mod_H))*(mod_SE/mod_I)
  I <- c(rep(0, mod_SE), rep(1, mod_I))
  
  t_samp <- sample_int_crank(length(I),
                             n_tests-round(mod_H),
                             prob = c(rep(1, mod_SE),
                                      rep(B, mod_I)))
  
  pct_pos <- sum(I[t_samp])/(n_tests-round(mod_H))
  
  return(pct_pos)
}

n_sim_tests <- 30

sim_tests_mat <- matrix(NA, ncol = n_sim_tests, 
                        nrow = nrow(sf_test_mod_join_samp_bias %>% 
                                      filter(Date >= as.Date("2020-03-15")) %>% 
                                      drop_na()))

set.seed(430)

for(i in 1:n_sim_tests){
  sim_tests_mat[,i] <- apply(sf_test_mod_join_samp_bias %>% 
                               filter(Date >= as.Date("2020-03-15")) %>% 
                               drop_na(),
                             1, function(t){
                               test_with_bias(as.numeric(t["pos"]), as.numeric(t["tests"]),
                                              as.numeric(t["I_tot"]), as.numeric(t["H_new"]), 
                                              as.numeric(t["mod_SE"]))
                             })
  print(i)
}

saveRDS(sim_tests_mat,
        "sim_test_pct_pos_6-15-20.rds")
```

```{r test_mod_pct_comp}
sim_tests_mat <- readRDS("sim_test_pct_pos_6-15-20.rds")

sim_tests_pos_med <- apply(sim_tests_mat, 1, median)
sim_tests_75th <- apply(sim_tests_mat, 1, quantile, 0.75)
sim_tests_25th <- apply(sim_tests_mat, 1, quantile, 0.25)

as.data.frame(cbind(sf_test_mod_join_samp_bias %>% 
                               filter(Date >= as.Date("2020-03-15")) %>% 
                               drop_na(),
                    sim_tests_pos_med, sim_tests_25th, sim_tests_75th)) %>% 
  ggplot() +
    theme_bw() +
    geom_point(aes(x = non_H_pos_rate, y = sim_tests_pos_med)) +
    geom_errorbar(aes(x = non_H_pos_rate, 
                      ymin = sim_tests_25th,
                      ymax = sim_tests_75th)) +
    geom_abline(intercept = 0, slope = 1, lty = 2) +
    labs(x = "Observed test positive rate minus H admissions",
         y = "Simulated test positive rate",
         title = "Comparison of simulated and observed test positive rates adjusting for bias")
```

This is ok as the most recent test positive rates (which are lower) are in the lower left of this plot where the simulation and observed match better. Higher observed test positive rates were earlier in the outbreak when less tests were avaialble and testing bias was probably even more sporadic.

## Model forecast  

Assuming testing bias and number of tests remain constant, now want to determine what testing (and hospitalizations since that is what model fitting is based on) may look like in the near future under different transmission scenarios. The model was fit through the first week of June (June 5th), and the different transmission scenarios assume:  

1. Transmission has remained constant with $\mathcal{R}_e\approx0.8$  
2. Transmission has increased above 1 as of June 15th with $\mathcal{R}_e\approx1.1$  
3. Transmission has increased well above 1 as of June 15th with $\mathcal{R}_e\approx1.4$  

```{r sim_extended_setup}
spline_ext <- splinefun(x = beta.time[,1],
                        y = as.numeric(log(c(best_pt_pars[2],best_pt_pars[2],
                                  as.numeric(best_pt_pars[5:8]), 
                                  get_beta(1.4, best_pt_pars[3], 
                                           best.pars[["gam_r"]], best.pars[["rho"]])))),
                        method = "natural")

smooth_ext <- approxfun(c(0:t.tot),
                        c(rep(log(best_pt_pars[2]),beta.time[2,1]), 
                          sapply(c(beta.time[2,1]+1):(t.tot+1),
                                 spline_ext)))

proj_tests_per_day = 2200
```

```{r sim.extended, eval = FALSE}
sim.ext <- as.data.frame(ode(y = init, 
                             times = 1:t.tot, 
                             parms = best.pars,
                             func = lemma_odes, 
                             fx = smooth_ext)) %>% 
  mutate(Date = as.Date(t0+time-1)) 

sim.ext$Re <- get_R(exp(sapply(1:t.tot, smooth_ext)),
                    best.pars[["alpha"]], best.pars[["gam_r"]], best.pars[["rho"]], 
                    sim.ext$S, best.pars[["N"]])

ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_line(data = sim.ext,
            aes(x = Date, y = H),
            col = "coral", size = 1.2) +
  geom_point(data = sf_all,
             aes(x = Date, y = HOSP_CONF+ICU_CONF),
             pch = 17) +
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = c(as.Date("2020-02-15"), t.end))

ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_line(data = sim.ext,
            aes(x = Date, y = D),
            col = "coral", size = 1.2) +
  geom_point(data = sf_all,
             aes(x = Date, y = cum_death),
             pch = 17) +
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = c(as.Date("2020-02-15"), t.end))

ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_line(data = sim.ext,
            aes(x = Date, y = Re),
            col = "coral", size = 1.2) +
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = c(as.Date("2020-02-15"), t.end)) +
  scale_y_continuous(breaks = seq(0,3.5,0.5),
                     limits = c(0,4))

```

```{r adaptau_mod}
library(adaptivetau)

# List of transitions
transitions = list(
  c(S = -1, E = 1),     #Susceptible becomes exposed
  c(E = -1, Ir = 1),    #Exposed becomes Infected, will recover
  c(E = -1, Ih = 1),    #Exposed becomes Infected, will be hospitalized
  c(Ir = -1, R = 1),    #Infected, mild recovers
  c(Ih = -1, H = 1),    #Infected severe becomes hospitalized
  c(H = -1, D = 1),     #Hospitalized, becomes Dead
  c(H = -1, R = 1))     #Hospitalized becomes recovered

# Rate simulator
rates_base <- function(x, p, t, fx) {
  S = x['S']
  E = x['E']
  Ir = x['Ir']
  Ih = x['Ih']
  H = x['H']
  R = x['R']
  D = x["D"]

  # Parameters
    N = p[['N']]             #population size
    beta =future_beta(t)     #transmission rate
    cr = p[['cr']]           #Relative contact rate between S and Ir
    ch = p[['ch']]           #Relative contact rate between S and Ih
    sigma = p[['sigma']]     #1/serial interval
    alpha = p[['alpha']]     #proportion severely symptomatic
    rho = p[['rho']]         #1/time between symptom onset and hospitalization
    gam_r = p[['gam_r']]     #1/time to recovery (non-infectiousness) for mild/asymptomatic
    gam_h = p[['gam_h']]     #1/time to removed (recovered or deceased) for hospitalized
    mu = p[['mu']]           #proportion of hospitalized cases who die
  
  return(c(S*beta*(Ir*cr+Ih*ch)/N, #Susceptible becomes exposed
           E*(1-alpha)*sigma,      #Exposed becomes Infected, pre-symptomatic
           E*alpha*sigma,          #Exposed becomes Infected, asymptomatic
           Ir*gam_r,               #Infected, pre-symptomatic becomes Infected, severe
           Ih*rho,                 #Infected, pre-symptomatic becomes Infected, mild
           H*mu*gam_h,             #Hospitalized becomes Dead
           H*(1-mu)*gam_h))        #Hospitalized becomes recovered
}

#Function to simulate and return dataframe
stoch.sim = function(init, trans, rates, pars, t_sim){
  
  ssa.sim <- adaptivetau::ssa.adaptivetau(init.values = init, 
                                          transitions = trans,
                                          rateFunc = rates,
                                          params = pars,
                                          tf=t_sim)
  
  return(ssa.sim)
}  

#Starting conditions from today
init_today <- as.numeric(round(best.sim[dim(best.sim)[1]-1, 2:8]))
names(init_today) <- colnames(best.sim[2:8])

```

```{r Hadm_mod}
# List of transitions dividing recovered between those that recover from H and Ir for identifiable stoichiometry
transitions_Hadm = list(
  c(S = -1, E = 1),     #Susceptible becomes exposed
  c(E = -1, Ir = 1),    #Exposed becomes Infected, will recover
  c(E = -1, Ih = 1),    #Exposed becomes Infected, will be hospitalized
  c(Ir = -1, Rr = 1),   #Infected, mild recovers
  c(Ih = -1, H = 1),    #Infected severe becomes hospitalized
  c(H = -1, D = 1),     #Hospitalized, becomes Dead
  c(H = -1, Rh = 1))     #Hospitalized becomes recovered

# Rate simulator
rates_Hadm <- function(x, p, t, fx) {
  S = x['S']
  E = x['E']
  Ir = x['Ir']
  Ih = x['Ih']
  H = x['H']
  Rr = x['Rr']
  Rh = x['Rh']
  D = x["D"]

  # Parameters
    N = p[['N']]             #population size
    beta =future_beta(t)     #transmission rate
    cr = p[['cr']]           #Relative contact rate between S and Ir
    ch = p[['ch']]           #Relative contact rate between S and Ih
    sigma = p[['sigma']]     #1/serial interval
    alpha = p[['alpha']]     #proportion severely symptomatic
    rho = p[['rho']]         #1/time between symptom onset and hospitalization
    gam_r = p[['gam_r']]     #1/time to recovery (non-infectiousness) for mild/asymptomatic
    gam_h = p[['gam_h']]     #1/time to removed (recovered or deceased) for hospitalized
    mu = p[['mu']]           #proportion of hospitalized cases who die
  
  return(c(S*beta*(Ir*cr+Ih*ch)/N, #Susceptible becomes exposed
           E*(1-alpha)*sigma,      #Exposed becomes Infected, pre-symptomatic
           E*alpha*sigma,          #Exposed becomes Infected, asymptomatic
           Ir*gam_r,               #Infected, pre-symptomatic becomes Infected, severe
           Ih*rho,                 #Infected, pre-symptomatic becomes Infected, mild
           H*mu*gam_h,             #Hospitalized becomes Dead
           H*(1-mu)*gam_h))        #Hospitalized becomes recovered
}

# Initial conditions with Rr and Rh
init_Hadm <- as.numeric(round(best.sim[dim(best.sim)[1]-1, 2:7]))
init_Hadm <- c(init_Hadm, as.numeric(round(best.sim[dim(best.sim)[1]-1, 8]*c(0.95,0.05))))

names(init_Hadm) <- c(colnames(best.sim[2:7]), "Rr", "Rh")

```

### $\mathcal{R}_e\rightarrow0.8$  

```{r stoch_sims_setup_Re08}
#Function for future beta conditions
t.future <- as.numeric(t.end - last_fit)

new.Re <- 0.8 
new.beta <- get_beta(new.Re, best.pars[["alpha"]], 
                     best.pars[["gam_r"]], best.pars[["rho"]])
t_til_new_beta <- 7
future_beta <- approxfun(c(0, t_til_new_beta,t.future), c(best_pt_pars[8], new.beta, new.beta), method = "linear")

```

```{r r08_sims, eval = FALSE}
set.seed(430)

future_runs_Hadm_Re08 <- bind_rows(lapply(1:100, function(i){
  as.data.frame(stoch.sim(init = init_Hadm, 
                          trans = transitions_Hadm, 
                          rates = rates_Hadm,
                          pars = best.pars,
                          t_sim = t.future)) %>% 
    mutate(sim = i,
           dt = time-lag(time),
           tests = round(dt*proj_tests_per_day),
           datetime = as.POSIXct(sf_test %>% 
                                   filter(Date == last_fit) %>% 
                                   pull(specimen_collection_date))+(time*86400))
  }))

saveRDS(future_runs_Hadm_Re08, 
        "future_stoch_runs_Re08_6-15-20.rds")

```

```{r r08_sims_plot, message = FALSE}
future_runs_Hadm_Re08 <- readRDS("future_stoch_runs_Re08_6-15-20.rds")

re08_stoch_sum <- future_runs_Hadm_Re08 %>% 
  arrange(time) %>% 
  mutate(time_cut = cut(time, breaks = c(0:t.end))) %>% 
  group_by(time_cut) %>% 
  summarise(S_med = median(S),
            S25th = quantile(S, 0.25),
            S75th = quantile(S, 0.75),
            E_med = median(E),
            E25th = quantile(E, 0.25),
            E75th = quantile(E, 0.75),
            Ia_med = median(Ir),
            Ir25th = quantile(Ir, 0.25),
            Ir75th = quantile(Ir, 0.75),
            Is_med = median(Ih),
            Ih25th = quantile(Ih, 0.25),
            Ih75th = quantile(Ih, 0.75),
            H_med = median(H),
            H25th = quantile(H, 0.25),
            H75th = quantile(H, 0.75),
            Rr_med = median(Rr),
            Rr25th = quantile(Rr, 0.25),
            Rr75th = quantile(Rr, 0.75),
            D_med = median(D),
            D25th = quantile(D, 0.25),
            D75th = quantile(D, 0.75),
            Rh_med = median(Rh),
            Rh25th = quantile(Rh, 0.25),
            Rh75th = quantile(Rh, 0.75),
            tests = sum(tests)/max(future_runs_Hadm_Re08$sim)) %>% 
  mutate(time = as.numeric(time_cut),
         Date = last_fit+time-2,
         Re = 0.8)

ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_col(data = sf_all,
           aes(x = Date, y = HOSP_CONF+ICU_CONF),
           col = "black", fill = "grey60")  +
  geom_line(data = best.sim,
            aes(x = Date, y = H),
            col = "coral", size = 1.2) +
  geom_ribbon(data = re08_stoch_sum,
              aes(x = Date, ymin = H25th, ymax = H75th),
              fill = "red", alpha = 0.3) +
  geom_line(data = re08_stoch_sum,
            aes(x = Date, y = H_med),
            size = 1.2, col = "red") +
  labs(x = "Date", y = "Hospitalizations",
       title = "Hospitalizations projection with Re->0.8")

```

```{r plot_stoch_sims_Hadm08, include = FALSE}
future_runs_Hadm_Re08 <- future_runs_Hadm_Re08 %>% 
  group_by(sim) %>% 
  mutate(dH = H - lag(H),
         dRh = Rh - lag(Rh),
         dD = D-lag(D),
         outH = dRh + dD,
         newH = dH+outH,
         tests_noH = tests-newH,
         tests_use = if_else(tests_noH < 1, NA_real_, tests_noH),
         Date = as.Date(datetime))

```

```{r future_test_pos_Re08, eval = FALSE}
#start <- Sys.time()
test_cl <- makeCluster(detectCores())         # Make cluster
clusterExport(test_cl, c("future_runs_Hadm_Re08", 
                         "bias_median"))         # Export objects to cluster

clusterEvalQ(test_cl, lapply(c("tidyverse", "wrswoR"), library, character.only = T))

stoch_sim_tests <- parApply(test_cl, future_runs_Hadm_Re08 %>% drop_na(), 1, 
                            function(t){
                              t <- as.numeric(t)
                              
                              inf_status <- c(rep(0, (t[2]+t[3])),    # Ss and Es
                                                 rep(1, (t[4]+t[5]))) # Irs and Ihs
                              samps <- t[12]-t[18]                  # number tests minus hospital admissions
                              probs <- c(rep(1, (t[2]+t[3])),
                                        rep(bias_median, (t[4]+t[5])))       # Sampling bias
                              
                              t_samp <- sample_int_crank(length(inf_status), samps, probs) 
                              
                              return(sum(inf_status[t_samp])/(t[12]-t[18])) #Return pct positive
                            })

stopCluster(test_cl)
#end <- Sys.time()

#as.numeric(end-start)*nrow(future_runs_Hadm)/360000

saveRDS(stoch_sim_tests, 
        "future_stoch_runs_tests_Re08_6-15-20.rds")

```

```{r plot_stoch_sims_test_r08}
stoch_sim_tests_Re08 <- readRDS("future_stoch_runs_tests_Re08_6-15-20.rds")

future_runs_tests_sum_Re08 <- as.data.frame(cbind(future_runs_Hadm_Re08 %>% drop_na(),
                                           sim_pct_pos = stoch_sim_tests_Re08)) %>% 
  arrange(time) %>% 
  mutate(time_cut = cut(time, breaks = c(0:t.future))) %>% 
  group_by(time_cut) %>% 
    summarise(pct_pos_med = median(sim_pct_pos),
              pct_pos_25th = quantile(sim_pct_pos, 0.25),
              pct_pos_75th = quantile(sim_pct_pos, 0.75)) %>% 
  mutate(time = as.numeric(time_cut),
         Date = last_fit+time-2,
         Re = 0.8)

  ggplot() +
    theme_bw() +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1)) +
  geom_line(data = sf_test %>% filter(Date >= as.Date("2020-04-23")),
            aes(x = Date, y = pct),
            size = 1.1)  +
  geom_ribbon(data = future_runs_tests_sum_Re08,
                aes(x = Date, ymin = pct_pos_25th, ymax = pct_pos_75th),
                alpha = 0.3, fill = "gold")+
  geom_line(data = future_runs_tests_sum_Re08,
            aes(x = Date, y = pct_pos_med),
            size = 1.2, col = "gold") +
  labs(x = "Date", y = "% Positive tests",
       title = "Future Positive test percentages",
       subtitle = paste("Assuming", proj_tests_per_day, "tests per day, B=", 
                        round(bias_median), ", Re->0.8"))

```

### $\mathcal{R}_e\rightarrow1.1$  

```{r stoch_sims_setup_Re11}
#Function for future beta conditions
t.future <- as.numeric(t.end - last_fit)

new.Re <- 1.1
new.beta <- get_beta(new.Re, best.pars[["alpha"]], 
                     best.pars[["gam_r"]], best.pars[["rho"]])
t_til_new_beta <- 7
future_beta <- approxfun(c(0, t_til_new_beta,t.future), c(best_pt_pars[8], new.beta, new.beta), method = "linear")

```

```{r r11_sims, eval = FALSE}
set.seed(430)

future_runs_Hadm_Re11 <- bind_rows(lapply(1:100, function(i){
  as.data.frame(stoch.sim(init = init_Hadm, 
                          trans = transitions_Hadm, 
                          rates = rates_Hadm,
                          pars = best.pars,
                          t_sim = t.future)) %>% 
    mutate(sim = i,
           dt = time-lag(time),
           tests = round(dt*proj_tests_per_day),
           datetime = as.POSIXct(sf_test %>% 
                                   filter(Date == last_fit) %>% 
                                   pull(specimen_collection_date))+(time*86400))
  }))

saveRDS(future_runs_Hadm_Re11, 
        "future_stoch_runs_Re11_6-15-20.rds")

```

```{r r11_sims_plot}
future_runs_Hadm_Re11 <- readRDS("future_stoch_runs_Re11_6-15-20.rds")

re11_stoch_sum <- future_runs_Hadm_Re11 %>% 
  arrange(time) %>% 
  mutate(time_cut = cut(time, breaks = c(0:t.end))) %>% 
  group_by(time_cut) %>% 
  summarise(S_med = median(S),
            S25th = quantile(S, 0.25),
            S75th = quantile(S, 0.75),
            E_med = median(E),
            E25th = quantile(E, 0.25),
            E75th = quantile(E, 0.75),
            Ia_med = median(Ir),
            Ir25th = quantile(Ir, 0.25),
            Ir75th = quantile(Ir, 0.75),
            Is_med = median(Ih),
            Ih25th = quantile(Ih, 0.25),
            Ih75th = quantile(Ih, 0.75),
            H_med = median(H),
            H25th = quantile(H, 0.25),
            H75th = quantile(H, 0.75),
            Rr_med = median(Rr),
            Rr25th = quantile(Rr, 0.25),
            Rr75th = quantile(Rr, 0.75),
            D_med = median(D),
            D25th = quantile(D, 0.25),
            D75th = quantile(D, 0.75),
            Rh_med = median(Rh),
            Rh25th = quantile(Rh, 0.25),
            Rh75th = quantile(Rh, 0.75),
            tests = sum(tests)/max(future_runs_Hadm_Re11$sim)) %>% 
  mutate(time = as.numeric(time_cut),
         Date = last_fit+time-2, 
         Re = 1.1)

ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_col(data = sf_all,
           aes(x = Date, y = HOSP_CONF+ICU_CONF),
           col = "black",  fill = "grey60")  +
  geom_line(data = best.sim,
            aes(x = Date, y = H),
            col = "coral", size = 1.2) +
  geom_ribbon(data = re11_stoch_sum,
              aes(x = Date, ymin = H25th, ymax = H75th),
              fill = "red", alpha = 0.3) +
  geom_line(data = re11_stoch_sum,
            aes(x = Date, y = H_med),
            size = 1.2, col = "red") +
  labs(x = "Date", y = "Hospitalizations",
       title = "Hospitalizations projection with Re->1.1")

```

```{r plot_stoch_sims_Hadm11, include = FALSE}
future_runs_Hadm_Re11 <- future_runs_Hadm_Re11 %>% 
  group_by(sim) %>% 
  mutate(dH = H - lag(H),
         dRh = Rh - lag(Rh),
         dD = D-lag(D),
         outH = dRh + dD,
         newH = dH+outH,
         tests_noH = tests-newH,
         tests_use = if_else(tests_noH < 1, NA_real_, tests_noH),
         Date = as.Date(datetime))
```

```{r future_test_pos_re11, eval = FALSE}
#start <- Sys.time()
test_cl <- makeCluster(detectCores())         # Make cluster
clusterExport(test_cl, c("future_runs_Hadm_Re11", 
                         "bias_median"))         # Export objects to cluster

clusterEvalQ(test_cl, lapply(c("tidyverse", "wrswoR"), library, character.only = T))

stoch_sim_tests <- parApply(test_cl, future_runs_Hadm_Re11 %>% drop_na(), 1, 
                            function(t){
                              t <- as.numeric(t)
                              
                              inf_status <- c(rep(0, (t[2]+t[3])),    # Ss and Es
                                                 rep(1, (t[4]+t[5]))) # Irs and Ihs
                              samps <- t[12]-t[18]                  # number tests minus hospital admissions
                              probs <- c(rep(1, (t[2]+t[3])),
                                        rep(bias_median, (t[4]+t[5])))       # Sampling bias
                              
                              t_samp <- sample_int_crank(length(inf_status), samps, probs) 
                              
                              return(sum(inf_status[t_samp])/(t[12]-t[18])) #Return pct positive
                            })

stopCluster(test_cl)
#end <- Sys.time()

#as.numeric(end-start)*nrow(future_runs_Hadm)/360000

saveRDS(stoch_sim_tests, 
        "future_stoch_runs_tests_Re11_6-15-20.rds")

```

```{r plot_stoch_sims_test_Re11}
stoch_sim_tests_Re11 <- readRDS("future_stoch_runs_tests_Re11_6-15-20.rds")

future_runs_tests_sum_Re11 <- as.data.frame(cbind(future_runs_Hadm_Re11 %>% drop_na(),
                                           sim_pct_pos = stoch_sim_tests_Re11)) %>% 
  arrange(time) %>% 
  mutate(time_cut = cut(time, breaks = c(0:t.future))) %>% 
  group_by(time_cut) %>% 
    summarise(pct_pos_med = median(sim_pct_pos),
              pct_pos_25th = quantile(sim_pct_pos, 0.25),
              pct_pos_75th = quantile(sim_pct_pos, 0.75)) %>% 
  mutate(time = as.numeric(time_cut),
         Date = last_fit+time-2,
         Re = 1.1)

  ggplot() +
    theme_bw() +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1)) +
  geom_line(data = sf_test %>% filter(Date >= as.Date("2020-04-23")),
            aes(x = Date, y = pct),
            size = 1.1)  +
  geom_ribbon(data = future_runs_tests_sum_Re11,
                aes(x = Date, ymin = pct_pos_25th, ymax = pct_pos_75th),
                alpha = 0.3, fill = "gold")+
  geom_line(data = future_runs_tests_sum_Re11,
            aes(x = Date, y = pct_pos_med),
            size = 1.2, col = "gold") +
  labs(x = "Date", y = "% Positive tests",
       title = "Future Positive test percentages",
       subtitle = paste("Assuming", proj_tests_per_day, "tests per day, B=",round(bias_median), ", Re->1.1"))

```


### $\mathcal{R}_e\rightarrow1.4$  

```{r stoch_sims_setup_Re14}
#Function for future beta conditions
t.future <- as.numeric(t.end - last_fit)

new.Re <- 1.4 
new.beta <- get_beta(new.Re, best.pars[["alpha"]], 
                     best.pars[["gam_r"]], best.pars[["rho"]])
t_til_new_beta <- 7
future_beta <- approxfun(c(0, t_til_new_beta,t.future), c(best_pt_pars[8], new.beta, new.beta), method = "linear")

```

```{r r14_sims}
set.seed(430)

future_runs_Hadm_Re14 <- bind_rows(lapply(1:100, function(i){
  as.data.frame(stoch.sim(init = init_Hadm, 
                          trans = transitions_Hadm, 
                          rates = rates_Hadm,
                          pars = best.pars,
                          t_sim = t.future)) %>% 
    mutate(sim = i,
           dt = time-lag(time),
           tests = round(dt*proj_tests_per_day),
           datetime = as.POSIXct(sf_test %>% 
                                   filter(Date == last_fit) %>% 
                                   pull(specimen_collection_date))+(time*86400))
  }))
saveRDS(future_runs_Hadm_Re14, 
        "future_stoch_runs_Re14_6-15-20.rds")

```

```{r r14_sims_plot}
future_runs_Hadm_Re14 <- readRDS("future_stoch_runs_Re14_6-15-20.rds")

re14_stoch_sum <- future_runs_Hadm_Re14 %>% 
  arrange(time) %>% 
  mutate(time_cut = cut(time, breaks = c(0:t.end))) %>% 
  group_by(time_cut) %>% 
  summarise(S_med = median(S),
            S25th = quantile(S, 0.25),
            S75th = quantile(S, 0.75),
            E_med = median(E),
            E25th = quantile(E, 0.25),
            E75th = quantile(E, 0.75),
            Ia_med = median(Ir),
            Ir25th = quantile(Ir, 0.25),
            Ir75th = quantile(Ir, 0.75),
            Is_med = median(Ih),
            Ih25th = quantile(Ih, 0.25),
            Ih75th = quantile(Ih, 0.75),
            H_med = median(H),
            H25th = quantile(H, 0.25),
            H75th = quantile(H, 0.75),
            Rr_med = median(Rr),
            Rr25th = quantile(Rr, 0.25),
            Rr75th = quantile(Rr, 0.75),
            D_med = median(D),
            D25th = quantile(D, 0.25),
            D75th = quantile(D, 0.75),
            Rh_med = median(Rh),
            Rh25th = quantile(Rh, 0.25),
            Rh75th = quantile(Rh, 0.75),
            tests = sum(tests)/max(future_runs_Hadm_Re14$sim)) %>% 
  mutate(time = as.numeric(time_cut),
         Date = last_fit+time-2, 
         Re = 1.4)



ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_col(data = sf_all,
           aes(x = Date, y = HOSP_CONF+ICU_CONF),
           fill = "grey60", col = "black")  +
  geom_line(data = best.sim,
            aes(x = Date, y = H),
            col = "coral", size = 1.2) +
  geom_ribbon(data = re14_stoch_sum,
              aes(x = Date, ymin = H25th, ymax = H75th),
              fill = "red", alpha = 0.3) +
  geom_line(data = re14_stoch_sum,
            aes(x = Date, y = H_med),
            size = 1.2, col = "red") +
  labs(x = "Date", y = "Hospitalizations",
       main = "Hospitalizations projection with Re->1.4")

```

```{r plot_stoch_sims_Hadm14, include = FALSE}
future_runs_Hadm_Re14 <- future_runs_Hadm_Re14 %>% 
  group_by(sim) %>% 
  mutate(dH = H - lag(H),
         dRh = Rh - lag(Rh),
         dD = D-lag(D),
         outH = dRh + dD,
         newH = dH+outH,
         tests_noH = tests-newH,
         tests_use = if_else(tests_noH < 1, NA_real_, tests_noH),
         Date = as.Date(datetime))

```

```{r future_test_pos_re14, eval = FALSE}
#start <- Sys.time()
test_cl <- makeCluster(detectCores())         # Make cluster
clusterExport(test_cl, c("future_runs_Hadm_Re14", 
                         "bias_median"))         # Export objects to cluster

clusterEvalQ(test_cl, lapply(c("tidyverse", "wrswoR"), library, character.only = T))

stoch_sim_tests <- parApply(test_cl, future_runs_Hadm_Re14 %>% drop_na(), 1, 
                            function(t){
                              t <- as.numeric(t)
                              
                              inf_status <- c(rep(0, (t[2]+t[3])),    # Ss and Es
                                                 rep(1, (t[4]+t[5]))) # Irs and Ihs
                              samps <- t[12]-t[18]                  # number tests minus hospital admissions
                              probs <- c(rep(1, (t[2]+t[3])),
                                        rep(bias_median, (t[4]+t[5])))       # Sampling bias
                              
                              t_samp <- sample_int_crank(length(inf_status), samps, probs) 
                              
                              return(sum(inf_status[t_samp])/(t[12]-t[18])) #Return pct positive
                            })

stopCluster(test_cl)
#end <- Sys.time()

#as.numeric(end-start)*nrow(future_runs_Hadm)/360000

saveRDS(stoch_sim_tests, 
        "future_stoch_runs_tests_Re14_6-15-20.rds")

```

```{r plot_stoch_sims_test_Re14}
stoch_sim_tests_Re14 <- readRDS("future_stoch_runs_tests_Re14_6-15-20.rds")

future_runs_tests_sum_Re14 <- as.data.frame(cbind(future_runs_Hadm_Re14 %>% drop_na(),
                                           sim_pct_pos = stoch_sim_tests_Re14)) %>% 
  arrange(time) %>% 
  mutate(time_cut = cut(time, breaks = c(0:t.future))) %>% 
  group_by(time_cut) %>% 
    summarise(pct_pos_med = median(sim_pct_pos),
              pct_pos_25th = quantile(sim_pct_pos, 0.25),
              pct_pos_75th = quantile(sim_pct_pos, 0.75)) %>% 
  mutate(time = as.numeric(time_cut),
         Date = last_fit+time-2,
         Re = 1.4)

  ggplot() +
    theme_bw() +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1)) +
  geom_line(data = sf_test %>% filter(Date >= as.Date("2020-04-23")),
            aes(x = Date, y = pct),
            size = 1.1)  +
  geom_ribbon(data = future_runs_tests_sum_Re14,
                aes(x = Date, ymin = pct_pos_25th, ymax = pct_pos_75th),
                alpha = 0.3, fill = "gold")+
  geom_line(data = future_runs_tests_sum_Re14,
            aes(x = Date, y = pct_pos_med),
            size = 1.2, col = "gold") +
  labs(x = "Date", y = "% Positive tests",
       title = "Future Positive test percentages",
       subtitle = paste("Assuming", proj_tests_per_day, "tests per day, B=", round(bias_median), ", Re->1.4"))

```

## Compare tests, hospitalizations in different future transmission scenarios  
```{r hosp_future}
all_stoch_sums <- as.data.frame(rbind(re08_stoch_sum,
                                      re11_stoch_sum,
                                      re14_stoch_sum)) 
  ggplot() +
    theme_bw() +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1),
          legend.text = element_text(size =14)) +
    geom_col(data = sf_all,
             aes(x = Date, y = HOSP_CONF+ICU_CONF),
             fill = "grey60", col = "black")  +
    geom_line(data = best.sim,
              aes(x = Date, y = H),
              col = "coral", size = 1.2) +
    geom_ribbon(data = all_stoch_sums,
                aes(x = Date, ymin = H25th, ymax = H75th,
                    fill = as.factor(Re)),
                alpha = 0.3) +
    geom_line(data = all_stoch_sums,
              aes(x =Date, y = H_med, col = as.factor(Re)),
              size = 1.2) +
    scale_fill_manual(values = c("gold1", "darkorange", "firebrick"),
                       labels = c("Re->0.8",
                                  "Re->1.1",
                                  "Re->1.4")) +
    scale_color_manual(values = c("gold1", "darkorange", "firebrick"),
                       labels = c("Re->0.8",
                                  "Re->1.1",
                                  "Re->1.4")) +
    labs(x = "Date", y = "Hospitalizations",
         title = "Future Hospitalizations projections under different transmission scenarios",
         fill = "", col = "")

```

```{r tests_future}
all_stoch_test_sums <- as.data.frame(rbind(future_runs_tests_sum_Re08,
                                           future_runs_tests_sum_Re11,
                                           future_runs_tests_sum_Re14)) 
  ggplot() +
    theme_bw() +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1),
          legend.text = element_text(size =14)) +
  geom_line(data = sf_test %>% filter(Date >= as.Date("2020-04-23")),
            aes(x = Date, y = pct),
            size = 1.1)  +
  geom_ribbon(data = all_stoch_test_sums,
                aes(x = Date, ymin = pct_pos_25th, ymax = pct_pos_75th, 
                    fill = as.factor(Re)),
                alpha = 0.3)+
  geom_line(data = all_stoch_test_sums,
            aes(x = Date, y = pct_pos_med,
                col = as.factor(Re)),
            size = 1.2) +
    scale_fill_manual(values = c("gold1", "darkorange", "firebrick"),
                       labels = c("Re->0.8",
                                  "Re->1.1",
                                  "Re->1.4")) +
    scale_color_manual(values = c("gold1", "darkorange", "firebrick"),
                       labels = c("Re->0.8",
                                  "Re->1.1",
                                  "Re->1.4")) +
  labs(x = "Date", y = "% Positive tests",
       title = "Future Positive test percentages under different transmission scenarios",
       subtitle = paste("Assuming", proj_tests_per_day, "tests per day, B=", round(bias_median), ", Re->0.8"),
       fill = "", col = "")

```

```{r pct_smooth}
test_pos_pct_smooth_plot <- ggplot() +
    theme_bw() +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1),
          legend.text = element_text(size =14)) +
  geom_line(data = sf_test %>% 
                mutate(pct_7dayavg = zoo::rollmean(pct, 7, fill = NA)) %>% 
                filter(Date >= as.Date("2020-04-23")),
              aes(x = Date, y = pct_7dayavg),
            size = 1.1)  +
  geom_ribbon(data = all_stoch_test_sums,
                aes(x = Date, ymin = pct_pos_25th, ymax = pct_pos_75th, 
                    fill = as.factor(Re)),
                alpha = 0.3)+
  geom_line(data = all_stoch_test_sums,
            aes(x = Date, y = pct_pos_med,
                col = as.factor(Re)),
            size = 1.2) +
    scale_fill_manual(values = c("gold1", "darkorange", "firebrick"),
                       labels = c("Re->0.8",
                                  "Re->1.1",
                                  "Re->1.4")) +
    scale_color_manual(values = c("gold1", "darkorange", "firebrick"),
                       labels = c("Re->0.8",
                                  "Re->1.1",
                                  "Re->1.4")) +
  labs(x = "Date", y = "% Positive tests (7-day average)",
       title = "Future Positive test percentages under different transmission scenarios",
       subtitle =paste("  Assuming ", proj_tests_per_day, " tests per day, B=", round(bias_median)),
       fill = "", col = "")

test_pos_pct_smooth_plot
```

```{r sf_tests_p100k}
  ggplot() +
    theme_bw() +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1),
          legend.position = "none") +
    geom_line(data = sf_test %>% 
                filter(Date >= as.Date("2020-04-23")),
              aes(x = Date, y = tests_p100k),
            size = 1.1)  +
    geom_line(data = all_stoch_sums %>% 
                group_by(Re) %>% 
                mutate(tests_smooth = rollmean(tests, 7, fill = NA, align = "right")),
              aes(x = Date, y = (tests_smooth/pop.size)*1e5, 
                  col = as.factor(Re)), size = 1.2) +
    scale_color_manual(values = c("gold1", "darkorange", "firebrick"),
                       labels = c("Re->0.8",
                                  "Re->1.1",
                                  "Re->1.4")) +
    labs(x = "Date",
         y = "Tests conducted per 100k",
         title = "Tests conducted leading up to and in simulations")
  
```

```{r sf_tests_p100k_smooth}
  ggplot() +
    theme_bw() +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1),
          legend.position = "none") +
    geom_line(data = sf_test %>% 
                filter(Date >= as.Date("2020-04-23")),
              aes(x = Date, y = tests_p100k_smooth),
            size = 1.1)  +
    geom_line(data = all_stoch_sums %>% 
                group_by(Re) %>% 
                mutate(tests_smooth = rollmean(tests, 7, fill = NA, align = "right")),
              aes(x = Date, y = (tests_smooth/pop.size)*1e5, 
                  col = as.factor(Re)), size = 1.2) +
    scale_color_manual(values = c("gold1", "darkorange", "firebrick"),
                       labels = c("Re->0.8",
                                  "Re->1.1",
                                  "Re->1.4")) +
    labs(x = "Date",
         y = "Tests conducted per 100k (7 day average)",
         title = "Tests conducted leading up to and in simulations")
  
```

## Consequences of unaccounted changes in sampling bias  
If sampling bias decreases such that infectious people are less likely to seek testing, our ability to detect an increase in $\mathcal{R}_e$ through testing data is likely to be limited, and may even give a false sense of security: positive % or number of positive tests may appear to be decreasing, even as transmission is increasing, if the sampling bias of infectious people has decreased. Therefore here the assumption that testing bias remains constant in the nearterm projections is changed to an assumption that the testing bias decreases by ~30%. This is feasible to imagine as people may become more complacent/accustomed to COVID and therefore decrease they're test-seeking behavior, especially among mild cases. E.g. someone with a headache and mild cough who would have been very concerned about COVID in April/May no longer seeks out testing or delays testing until more serious symptoms arise or persist.

```{r bias_reduction}
bias_reduced <- floor(bias_median*0.7)

```

```{r future_test_pos_re08_bias_decrease, eval = FALSE}
bias_reduced <- floor(bias_median*0.7)

#start <- Sys.time()
test_cl <- makeCluster(detectCores())         # Make cluster
clusterExport(test_cl, c("future_runs_Hadm_Re08", 
                         "bias_reduced"))         # Export objects to cluster

clusterEvalQ(test_cl, lapply(c("tidyverse", "wrswoR"), library, character.only = T))

stoch_sim_tests <- parApply(test_cl, future_runs_Hadm_Re08 %>% drop_na(), 1, 
                            function(t){
                              t <- as.numeric(t)
                              
                              inf_status <- c(rep(0, (t[2]+t[3])),    # Ss and Es
                                                 rep(1, (t[4]+t[5]))) # Irs and Ihs
                              samps <- t[12]-t[18]                  # number tests minus hospital admissions
                              probs <- c(rep(1, (t[2]+t[3])),
                                        rep(bias_reduced, (t[4]+t[5])))       # Sampling bias
                              
                              t_samp <- sample_int_crank(length(inf_status), samps, probs) 
                              
                              return(sum(inf_status[t_samp])/(t[12]-t[18])) #Return pct positive
                            })

stopCluster(test_cl)
#end <- Sys.time()

#as.numeric(end-start)*nrow(future_runs_Hadm)/360000

saveRDS(stoch_sim_tests, 
        "future_stoch_runs_tests_Re08_decrease_bias_6-15-20.rds")

```

```{r future_test_pos_re11_bias_decrease, eval = FALSE}
#start <- Sys.time()
test_cl <- makeCluster(detectCores())         # Make cluster
clusterExport(test_cl, c("future_runs_Hadm_Re11", 
                         "bias_reduced"))         # Export objects to cluster

clusterEvalQ(test_cl, lapply(c("tidyverse", "wrswoR"), library, character.only = T))

stoch_sim_tests <- parApply(test_cl, future_runs_Hadm_Re11 %>% drop_na(), 1, 
                            function(t){
                              t <- as.numeric(t)
                              
                              inf_status <- c(rep(0, (t[2]+t[3])),    # Ss and Es
                                                 rep(1, (t[4]+t[5]))) # Irs and Ihs
                              samps <- t[12]-t[18]                  # number tests minus hospital admissions
                              probs <- c(rep(1, (t[2]+t[3])),
                                        rep(bias_reduced, (t[4]+t[5])))       # Sampling bias
                              
                              t_samp <- sample_int_crank(length(inf_status), samps, probs) 
                              
                              return(sum(inf_status[t_samp])/(t[12]-t[18])) #Return pct positive
                            })

stopCluster(test_cl)
#end <- Sys.time()

#as.numeric(end-start)*nrow(future_runs_Hadm)/360000

saveRDS(stoch_sim_tests, 
        "future_stoch_runs_tests_Re11_decrease_bias_6-15-20.rds")

```

```{r future_test_pos_re14_bias_decrease, eval = FALSE}
#start <- Sys.time()
test_cl <- makeCluster(detectCores())         # Make cluster
clusterExport(test_cl, c("future_runs_Hadm_Re14", 
                         "bias_reduced"))         # Export objects to cluster

clusterEvalQ(test_cl, lapply(c("tidyverse", "wrswoR"), library, character.only = T))

stoch_sim_tests <- parApply(test_cl, future_runs_Hadm_Re14 %>% drop_na(), 1, 
                            function(t){
                              t <- as.numeric(t)
                              
                              inf_status <- c(rep(0, (t[2]+t[3])),    # Ss and Es
                                                 rep(1, (t[4]+t[5]))) # Irs and Ihs
                              samps <- t[12]-t[18]                  # number tests minus hospital admissions
                              probs <- c(rep(1, (t[2]+t[3])),
                                        rep(bias_reduced, (t[4]+t[5])))       # Sampling bias
                              
                              t_samp <- sample_int_crank(length(inf_status), samps, probs) 
                              
                              return(sum(inf_status[t_samp])/(t[12]-t[18])) #Return pct positive
                            })

stopCluster(test_cl)
#end <- Sys.time()

#as.numeric(end-start)*nrow(future_runs_Hadm)/360000

saveRDS(stoch_sim_tests, 
        "future_stoch_runs_tests_Re14_decrease_bias_6-15-20.rds")

```

```{r bias_decrease_sums}
# Re=0.8, bias reduced by ~30%

stoch_sim_tests_decrease_bias_Re08 <- readRDS("future_stoch_runs_tests_Re08_decrease_bias_6-15-20.rds")

future_runs_tests_decrease_bias_sum_Re08 <- as.data.frame(cbind(future_runs_Hadm_Re08 %>% drop_na(),
                                                                sim_pct_pos = stoch_sim_tests_decrease_bias_Re08)) %>% 
  arrange(time) %>% 
  mutate(time_cut = cut(time, breaks = c(0:t.future))) %>% 
  group_by(time_cut) %>% 
  summarise(pct_pos_med = median(sim_pct_pos),
            pct_pos_25th = quantile(sim_pct_pos, 0.25),
            pct_pos_75th = quantile(sim_pct_pos, 0.75)) %>% 
  mutate(time = as.numeric(time_cut),
         Date = last_fit+time-2,
         Re = 0.8)

# Re=0.8, bias reduced by ~30%

stoch_sim_tests_decrease_bias_Re11 <- readRDS("future_stoch_runs_tests_Re11_decrease_bias_6-15-20.rds")

future_runs_tests_decrease_bias_sum_Re11 <- as.data.frame(cbind(future_runs_Hadm_Re11 %>% drop_na(),
                                                                sim_pct_pos = stoch_sim_tests_decrease_bias_Re11)) %>% 
  arrange(time) %>% 
  mutate(time_cut = cut(time, breaks = c(0:t.future))) %>% 
  group_by(time_cut) %>% 
  summarise(pct_pos_med = median(sim_pct_pos),
            pct_pos_25th = quantile(sim_pct_pos, 0.25),
            pct_pos_75th = quantile(sim_pct_pos, 0.75)) %>% 
  mutate(time = as.numeric(time_cut),
         Date = last_fit+time-2,
         Re = 1.1)

# Re=0.8, bias reduced by ~30%

stoch_sim_tests_decrease_bias_Re14 <- readRDS("future_stoch_runs_tests_Re14_decrease_bias_6-15-20.rds")

future_runs_tests_decrease_bias_sum_Re14 <- as.data.frame(cbind(future_runs_Hadm_Re14 %>% drop_na(),
                                                                sim_pct_pos = stoch_sim_tests_decrease_bias_Re14)) %>% 
  arrange(time) %>% 
  mutate(time_cut = cut(time, breaks = c(0:t.future))) %>% 
  group_by(time_cut) %>% 
  summarise(pct_pos_med = median(sim_pct_pos),
            pct_pos_25th = quantile(sim_pct_pos, 0.25),
            pct_pos_75th = quantile(sim_pct_pos, 0.75)) %>% 
  mutate(time = as.numeric(time_cut),
         Date = last_fit+time-2,
         Re = 1.4)

```

```{r tests_future_decrease_bias}
all_stoch_test_sums_decrease_bias <- as.data.frame(rbind(future_runs_tests_decrease_bias_sum_Re08,
                                                         future_runs_tests_decrease_bias_sum_Re11,
                                                         future_runs_tests_decrease_bias_sum_Re14)) 
  ggplot() +
    theme_bw() +
    theme(axis.title = element_text(size = 14,
                                    face = "bold"),
          axis.text = element_text(size = 12),
          axis.text.x = element_text(angle = 45,
                                     hjust = 1),
          legend.text = element_text(size =14)) +
  geom_line(data = sf_test %>% 
                mutate(pct_7dayavg = zoo::rollmean(pct, 7, fill = NA)) %>% 
                filter(Date >= as.Date("2020-04-23")),
              aes(x = Date, y = pct_7dayavg),
            size = 1.1)  +
  geom_ribbon(data = all_stoch_test_sums_decrease_bias,
                aes(x = Date, ymin = pct_pos_25th, ymax = pct_pos_75th, 
                    fill = as.factor(Re)),
                alpha = 0.3)+
  geom_line(data = all_stoch_test_sums_decrease_bias,
            aes(x = Date, y = pct_pos_med,
                col = as.factor(Re)),
            size = 1.2, lty = 2) +
    scale_fill_manual(values = c("gold1", "darkorange", "firebrick"),
                       labels = c("Re->0.8",
                                  "Re->1.1",
                                  "Re->1.4")) +
    scale_color_manual(values = c("gold1", "darkorange", "firebrick"),
                       labels = c("Re->0.8",
                                  "Re->1.1",
                                  "Re->1.4")) +
  labs(x = "Date", y = "% Positive tests (7-day average)",
       title = "Future Positive test percentages under different transmission scenarios",
       subtitle = paste("  Assuming", proj_tests_per_day, "tests per day, B=", round(bias_reduced)),
       fill = "", col = "")

```

```{r comp_bias_reduce}
all_stoch_test_sums2 <- all_stoch_test_sums %>% 
  mutate(bias = "Estimated")

all_stoch_test_sums_decrease_bias2 <- all_stoch_test_sums_decrease_bias %>% 
  mutate(bias = "Reduced")

all_stoch_test_sums_bias_and_normal <- rbind(all_stoch_test_sums2,
                                             all_stoch_test_sums_decrease_bias2)

ggplot() +
  theme_bw() +
  theme(axis.title = element_text(size = 14,
                                  face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        legend.text = element_text(size =14)) +
  geom_line(data = sf_test %>% 
              mutate(pct_7dayavg = zoo::rollmean(pct, 7, fill = NA)) %>% 
              filter(Date >= as.Date("2020-04-23")),
            aes(x = Date, y = pct_7dayavg),
            size = 1.1)  +
  geom_line(data = all_stoch_test_sums_bias_and_normal,
            aes(x = Date, y = pct_pos_med,
                col = as.factor(Re), lty = bias),
            size = 1.2) +
  scale_color_manual(values = c("gold1", "darkorange", "firebrick"),
                     labels = c("Re->0.8",
                                "Re->1.1",
                                "Re->1.4")) +   
  labs(x = "Date", y = "% Positive tests (7-day average)",
       title = "Effect of reduced bias on future positive test percentage",
       fill = "", col = "")


```

