---
title: "Stan fit StL and TN"
author: "Chris Hoover"
date: "6/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.align = "left",
                      fig.height = 6,
                      fig.width = 8)

require(rstan)
require(readxl)
require(tidyverse)

# Rstan prep
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')

# Summary function to get summaries across posteriors
stan_post_sum <- function(d){
  thing.med <- median(d)
  thing.mean <- mean(d)
  thing.75 <- quantile(d, 0.75)
  thing.25 <- quantile(d, 0.25)
  thing.975 <- quantile(d, 0.975)
  thing.025 <- quantile(d, 0.025)
  
  return(c("q.mean" = thing.mean, 
           "q.50" = thing.med, 
           "q.75" = thing.75, 
           "q.25" = thing.25, 
           "q.975" = thing.975, 
           "q.025" = thing.025))
}

```

```{r fit_data}
TN_all <- readRDS("../TN/Data/TN_all_2020-06-04.RDS")
TN_nsh <- readRDS("../TN/Data/TN_counties_2020-06-01.RDS") %>% 
  filter(COUNTY == "Davidson")
StL <- read_xlsx("Data/StLMay20-adjusted priors.xlsx", sheet = "Data",
                 skip = 8)

```

```{r mod_setup}
start.date <- as.Date("2020-01-23")
end.date <- as.Date("2020-07-01")

stl_inputs <- list()

# the number of age groups
  stl_inputs[['nage']] = 1 

# the population for each age group
  stl_inputs[['npop']] = array(400000)
# the fraction of hospitalized cases (ICU + non-ICU)
  stl_inputs[['frac_hosp']] = array(0.07)
# the fraction of ICU cases
  stl_inputs[['frac_icu']] = array(0.02)
# the death rate of infected
  stl_inputs[['frac_mort']] = array(0.01)
# the fraction of asymptomatic cases
  stl_inputs[['frac_asym']] = array(0.178)

# a parameter modifying the prior uncertainty in the fractions above (set to 1000.0 for `nage` > 1)
  stl_inputs[['alpha_multiplier']] = 100.0

# mean duration in "exposed" stage
  stl_inputs[['mu_duration_lat']] = 5.0
# standard deviation (sd) of duration in "exposed" stage
  stl_inputs[['sigma_duration_lat']] = 2.0
# mean duration in "infectious" stage for asymptomatic cases
  stl_inputs[['mu_duration_rec_asym']] = 7.0
# sd of duration in "infectious" stage for asymptomatic cases
  stl_inputs[['sigma_duration_rec_asym']] = 5.0
# mean duration in "infectious" stage for mild cases
  stl_inputs[['mu_duration_rec_mild']] = 7.0
# sd of duration in "infectious" stage for mild cases
  stl_inputs[['sigma_duration_rec_mild']] = 5.0
# mean duration in "infectious" stage for hospitalized cases
  stl_inputs[['mu_duration_pre_hosp']] = 5.0
# sd of duration in "infectious" stage for hospitalized cases
  stl_inputs[['sigma_duration_pre_hosp']] = 1.0
# mean duration in hospital for non-ICU cases
  stl_inputs[['mu_duration_hosp_mod']] = 12.0
# sd of duration in hospital for non-ICU cases
  stl_inputs[['sigma_duration_hosp_mod']] = 2.0
# mean duration in hospital for ICU cases
  stl_inputs[['mu_duration_hosp_icu']] = 12.0
# sd of duration in hospital for ICU cases
  stl_inputs[['sigma_duration_hosp_icu']] = 2.0

# lambda parameter for initial conditions of "exposed"
  stl_inputs[['lambda_ini_exposed']] = 0.3

# mean initial beta estimate
  stl_inputs[['mu_beta1']] = 0.2
# sd initial beta estimate
  stl_inputs[['sigma_beta1']] = 0.02

# distance in time between the knots used to construct the splines 
  stl_inputs[['dknot']] = 10

# spline mode (must be 1 or 2)
# splinemode = 1: estimate beta up to today
# splinemode = 2: estimate beta up to dknot days before today, then assume constant value up to today
  stl_inputs[['splinemode']] = 1

# number of interventions
  stl_inputs[['ninter']] = 1
  
# length of each intervention
  stl_inputs[['len_inter']] = array(10)
# mean change in beta through intervention 
  stl_inputs[['mu_beta_inter']] = array(1.0)
# sd change in beta through intervention
  stl_inputs[['sigma_beta_inter']] = array(0.5)

tn_inputs <- stl_inputs
nsh_inputs <- stl_inputs

#Pop of TN
tn_inputs[["npop"]] <- array(6829174)

#Pop of Davidson county
nsh_inputs[["npop"]] <- array(694144)
```

# St. Louis  
## Fit to StL hospital census data  

```{r stl_run, cache = T}
# Timeframe for simulation. Can replace Sys.Date() with any date at which to cut off model fitting
  itoday_1based = Sys.Date()-start.date
  
# specify the date up to when beta is estimated
  stl_inputs[['itoday']] = as.numeric(itoday_1based)

# the number of days to run the model for
  stl_inputs[['nt']] = as.numeric(as.Date(end.date) - as.Date(start.date))
  
# Put data into list. Added some leading 0s to help fitting
  stl_inputs[['nhobs']] <- nrow(StL)+10
  stl_inputs[['thobs']] <- as.numeric(c(as.Date(StL$date[1]) - start.date+c(-10:-1),
                                        as.Date(StL$date) - start.date))
  stl_inputs[['obs_Hmod']] <- c(rep(0,10), StL$hosp.lower/0.9)

# start time of each interventions
  stl_inputs[['t_inter']] = array(as.numeric(itoday_1based+5))

# Run the stan model  
StL_fit <- stan(
  file = "SEIR_Hosp_Only.stan",
  data = stl_inputs,
  thin = 5,
  chains = 4,
  warmup = 500,
  iter = 3000,
  cores = 4,
  refresh = 100,
  control = list(adapt_delta = 0.95)
)

stan_StL <- rstan::extract(StL_fit, permuted = T)

# State variables summary from model
state_vars <- c("S", "E", "Ia", "Im", "Ip", "Hosp", "Hicu", "Rlive", "Rdead")
  
StL_state_vars <- do.call(rbind, lapply(1:dim(stan_StL$x)[2], function(s){
  as_tibble(t(apply(stan_StL$x[,s,],2,stan_post_sum))) %>% 
    mutate(State = state_vars[s],
           obs = row_number(),
           Date = as.Date(start.date+obs),
           fit = "Hospital Census")
}))

colnames(StL_state_vars)[1:6] <- c("var_mean", "var_med", "var_75th", "var_25th", "var_975th", "var_025th")

# Hospitalizations output
StL_hosp <- stan_StL$hospitalized

StL_hosp_sum <- t(apply(StL_hosp, 2, stan_post_sum))

colnames(StL_hosp_sum) <- c("hosp_mean", "hosp_med", "hosp_75th", "hosp_25th", "hosp_975th", "hosp_025th")
 
StL_hosp_sum <- StL_hosp_sum %>% as_tibble() %>% 
  mutate(obs = row_number(),
         Date = as.Date(start.date+obs))

```

### Diagnostics  
#### Traceplots 
```{r stl_traces}
traceplot(StL_fit,
          pars = names(stan_StL)[which(grepl("duration", names(stan_StL)))])
```

#### Posteriors  
```{r stl_posteriors}
stan_dens(StL_fit, 
          pars = names(stan_StL)[which(grepl("duration", names(stan_StL)))], 
          separate_chains = TRUE)

```

#### Comparison to observed data    
```{r StL_plot}
ggplot() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11, 
                                   angle = 45,
                                   hjust = 1),
        axis.title = element_text(size = 14)) +
# Hospitalizations uncertainty intervals  
  geom_ribbon(data = StL_hosp_sum,
              aes(x = Date, 
                  ymax = hosp_75th,
                  ymin = hosp_25th),
              fill = "darkblue",
              alpha = 0.3) +
  geom_ribbon(data = StL_hosp_sum, 
              aes(x = Date, 
                  ymax = hosp_975th,
                  ymin = hosp_025th),
              fill = "darkblue",
              alpha = 0.1) +
# Observed data
  geom_point(data = StL,
             aes(x = as.Date(date), y = hosp.lower/0.9),
             pch = 17) +
# Model means
  geom_line(data = StL_hosp_sum,
            aes(x = Date, y = hosp_med),
            col = "darkblue") +
# Formatting
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = as.Date(c("2020-02-16",
                                  "2020-07-01")))+
  ylim(c(0, max(StL$hosp.lower/0.9, na.rm = T)+50)) +
  labs(y = "Hospitalizations",
       title = "StL hospital census fit")

```

## Fit to StL hospital census and admissions data  

```{r stl_run2, cache = T}
stl_inputs2 <- stl_inputs

# Add admissions data with some leading 0s
  stl_inputs2[['obs_Hadm']] <- c(rep(0,10),StL$new.admits.lower/0.9)

# Run the stan model  
StL_fit2 <- stan(
  file = "SEIR_Hosp_Census_Admissions.stan",
  data = stl_inputs2,
  thin = 5,
  chains = 4,
  warmup = 500,
  iter = 3000,
  cores = 4,
  refresh = 100,
  control = list(adapt_delta = 0.95)
)

stan_StL2 <- rstan::extract(StL_fit2, permuted = T)

# State variables summary from model
StL_state_vars2 <- do.call(rbind, lapply(1:dim(stan_StL2$x)[2], function(s){
  as_tibble(t(apply(stan_StL2$x[,s,],2,stan_post_sum))) %>% 
    mutate(State = state_vars[s],
           obs = row_number(),
           Date = as.Date(start.date+obs),
           fit = "Hospital Census + Admissions")
}))

colnames(StL_state_vars2)[1:6] <- c("var_mean", "var_med", "var_75th", "var_25th", "var_975th", "var_025th")

# Hospitalizations census output
StL_hosp2 <- stan_StL2$hospitalized

StL_hosp_sum2 <- t(apply(StL_hosp2, 2, stan_post_sum))

colnames(StL_hosp_sum2) <- c("hosp_mean", "hosp_med", "hosp_75th", "hosp_25th", "hosp_975th", "hosp_025th")
 
StL_hosp_sum2 <- StL_hosp_sum2 %>% as_tibble() %>% 
  mutate(obs = row_number(),
         Date = as.Date(start.date+obs))

# Hospitalizations admissions output
StL_adm2 <- stan_StL2$hosp_adm

StL_adm_sum2 <- t(apply(StL_adm2, 2, stan_post_sum))

colnames(StL_adm_sum2) <- c("adm_mean", "adm_med", "adm_75th", "adm_25th", "adm_975th", "adm_025th")
 
StL_adm_sum2 <- StL_adm_sum2 %>% as_tibble() %>% 
  mutate(obs = row_number(),
         Date = as.Date(start.date+obs))

```

### Diagnostics  
#### Traceplots 
```{r stl_traces2}
traceplot(StL_fit2,
          pars = names(stan_StL2)[which(grepl("duration", names(stan_StL2)))])
```

#### Posteriors  
```{r stl_posteriors2}
stan_dens(StL_fit2, 
          pars = names(stan_StL2)[which(grepl("duration", names(stan_StL2)))], 
          separate_chains = TRUE)

```

#### Comparison to observed data    
```{r StL_plot2}
ggplot() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11, 
                                   angle = 45,
                                   hjust = 1),
        axis.title = element_text(size = 14)) +
# Hospitalizations uncertainty intervals  
  geom_ribbon(data = StL_hosp_sum2,
              aes(x = Date, 
                  ymax = hosp_75th,
                  ymin = hosp_25th),
              fill = "darkblue",
              alpha = 0.3) +
  geom_ribbon(data = StL_hosp_sum2, 
              aes(x = Date, 
                  ymax = hosp_975th,
                  ymin = hosp_025th),
              fill = "darkblue",
              alpha = 0.1) +
# Observed data
  geom_point(data = StL,
             aes(x = as.Date(date), y = hosp.lower/0.9),
             pch = 17) +
# Model means
  geom_line(data = StL_hosp_sum2,
            aes(x = Date, y = hosp_med),
            col = "darkblue") +
# Formatting
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = as.Date(c("2020-02-16",
                                  "2020-07-01")))+
  ylim(c(0, max(StL$hosp.lower/0.9, na.rm = T)+50)) +
  labs(y = "Hospitalizations",
       title = "StL hospital admissions + census fit")

```

```{r StL_adm_plot2}
ggplot() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11, 
                                   angle = 45,
                                   hjust = 1),
        axis.title = element_text(size = 14)) +
# Hospitalizations uncertainty intervals  
  geom_ribbon(data = StL_adm_sum2,
              aes(x = Date, 
                  ymax = adm_75th,
                  ymin = adm_25th),
              fill = "darkblue",
              alpha = 0.3) +
  geom_ribbon(data = StL_adm_sum2, 
              aes(x = Date, 
                  ymax = adm_975th,
                  ymin = adm_025th),
              fill = "darkblue",
              alpha = 0.1) +
# Observed data
  geom_point(data = StL,
             aes(x = as.Date(date), y = new.admits.lower/0.9),
             pch = 17) +
# Model means
  geom_line(data = StL_adm_sum2,
            aes(x = Date, y = adm_med),
            col = "darkblue") +
# Formatting
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = as.Date(c("2020-02-16",
                                  "2020-07-01")))+
  ylim(c(0, max(StL$hosp.lower/0.9, na.rm = T)+50)) +
  labs(y = "Hospital admissions",
       title = "StL hospital admissions + census fit")

```


## Fit to StL hospital admissions data only  

```{r stl_run3, cache = T}
# Can use same inputs for census plus admits, it will just have some extra data  
# Run the stan model  
StL_fit3 <- stan(
  file = "SEIR_Hosp_Admissions.stan",
  data = stl_inputs2,
  thin = 5,
  chains = 4,
  warmup = 500,
  iter = 3000,
  cores = 4,
  refresh = 100,
  control = list(adapt_delta = 0.95)
)

stan_StL3 <- rstan::extract(StL_fit3, permuted = T)

# State variables summary from model
StL_state_vars3 <- do.call(rbind, lapply(1:dim(stan_StL3$x)[2], function(s){
  as_tibble(t(apply(stan_StL3$x[,s,],2,stan_post_sum))) %>% 
    mutate(State = state_vars[s],
           obs = row_number(),
           Date = as.Date(start.date+obs),
           fit = "Hospital Admissions")
}))

colnames(StL_state_vars3)[1:6] <- c("var_mean", "var_med", "var_75th", "var_25th", "var_975th", "var_025th")

# Hospitalizations admissions output
StL_adm3 <- stan_StL3$hosp_adm

StL_adm_sum3 <- t(apply(StL_adm3, 2, stan_post_sum))

colnames(StL_adm_sum3) <- c("adm_mean", "adm_med", "adm_75th", "adm_25th", "adm_975th", "adm_025th")
 
StL_adm_sum3 <- StL_adm_sum3 %>% as_tibble() %>% 
  mutate(obs = row_number(),
         Date = as.Date(start.date+obs))

```

### Diagnostics  
#### Traceplots 
```{r stl_traces3}
traceplot(StL_fit3,
          pars = names(stan_StL3)[which(grepl("duration", names(stan_StL3)))])
```

#### Posteriors  
```{r stl_posteriors3}
stan_dens(StL_fit3, 
          pars = names(stan_StL3)[which(grepl("duration", names(stan_StL3)))], 
          separate_chains = TRUE)

```

#### Comparison to observed data    

```{r StL_adm_plot3}
ggplot() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11, 
                                   angle = 45,
                                   hjust = 1),
        axis.title = element_text(size = 14)) +
# Hospitalizations uncertainty intervals  
  geom_ribbon(data = StL_adm_sum3,
              aes(x = Date, 
                  ymax = adm_75th,
                  ymin = adm_25th),
              fill = "darkblue",
              alpha = 0.3) +
  geom_ribbon(data = StL_adm_sum3, 
              aes(x = Date, 
                  ymax = adm_975th,
                  ymin = adm_025th),
              fill = "darkblue",
              alpha = 0.1) +
# Observed data
  geom_point(data = StL,
             aes(x = as.Date(date), y = new.admits.lower/0.9),
             pch = 17) +
# Model means
  geom_line(data = StL_adm_sum3,
            aes(x = Date, y = adm_med),
            col = "darkblue") +
# Formatting
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = as.Date(c("2020-02-16",
                                  "2020-07-01")))+
  ylim(c(0, max(StL$hosp.lower/0.9, na.rm = T)+50)) +
  labs(y = "Hospital Admissions",
       title = "StL hospital admissions fit")

```

## Compare StL fits across different data sources used for fitting  
```{r plot_comp}
as_tibble(rbind(StL_state_vars, StL_state_vars2, StL_state_vars3)) %>% 
  filter(State %in% c("Hosp", "Hicu", "Rdead")) %>% 
  ggplot() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 11, 
                                     angle = 45,
                                     hjust = 1),
          axis.title = element_text(size = 14)) +
  # Hospitalizations uncertainty intervals  
    geom_ribbon(aes(x = Date, 
                    ymax = var_975th,
                    ymin = var_025th,
                    fill = fit),
                alpha = 0.1) +
    geom_ribbon(aes(x = Date, 
                    ymax = var_75th,
                    ymin = var_25th,
                    fill = fit),
                alpha = 0.3) +
  # Model means
    geom_line(aes(x = Date, y = var_med, col = fit)) +
  # Formatting
    facet_wrap(State~., nrow = 3, scales = "free") +
    scale_x_date(date_labels = "%m/%d", 
                 date_breaks = "7 day",
                 limits = as.Date(c("2020-02-16",
                                    "2020-07-01")))+
    labs(y = "Hospital Admissions",
         title = "StL hospital admissions fit")

```

```{r plot_sigma_Hmod}
StL_sigma_Hmods <- data.frame("sigma_Hmod" = c(stan_StL$sigma_Hmod, stan_StL2$sigma_Hmod, stan_StL3$sigma_Hmod),
                              "fit" = c(rep("Hospital Census", length(stan_StL$sigma_Hmod)),
                                        rep("Hospital Census + Admissions", length(stan_StL2$sigma_Hmod)),
                                        rep("Hospital Admissions", length(stan_StL3$sigma_Hmod))))

StL_sigma_Hmods %>% 
  ggplot() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 11),
          axis.title = element_text(size = 14)) +
    geom_density(aes(sigma_Hmod, fill = fit),
                 alpha = 0.6)
```

# Tennessee  
## Fit to TN hospital admissions only   

```{r TN_adm, cache = T}
# specify the date up to when beta is estimated
  tn_inputs[['itoday']] = as.numeric(itoday_1based)

# the number of days to run the model for
  tn_inputs[['nt']] = as.numeric(as.Date(end.date) - as.Date(start.date))
  
# Put data into list
  tn_inputs[['nhobs']] <- nrow(TN_all %>% filter(!is.na(NEW_HOSP))) + 10
  
  TN_start <- as.Date(TN_all %>% 
                        filter(!is.na(NEW_HOSP)) %>% 
                        pull(DATE))[1]
  
  tn_inputs[['thobs']] <- as.numeric(c(TN_start - start.date + c(-10:-1),
                                       as.Date(TN_all %>% 
                                               filter(!is.na(NEW_HOSP)) %>% 
                                               pull(DATE)) - start.date))
  
  tn_inputs[['obs_Hadm']] <- c(rep(0,10), 
                               TN_all %>% filter(!is.na(NEW_HOSP)) %>% pull(NEW_HOSP))

# start time of each interventions
  tn_inputs[['t_inter']] = array(as.numeric(itoday_1based+5))
  
# Run the stan model  
TN_fit <- stan(
  file = "SEIR_Hosp_Admissions.stan",
  data = tn_inputs,
  thin = 5,
  chains = 4,
  warmup = 500,
  iter = 3000,
  cores = 4,
  refresh = 100,
  control = list(adapt_delta = 0.95)
)

stan_TN_adm <- rstan::extract(TN_fit, permuted = T)

# Hospitalizations admissions output
TN_adm <- stan_TN_adm$hosp_adm

TN_adm_sum <- t(apply(TN_adm, 2, stan_post_sum))

colnames(TN_adm_sum) <- c("adm_mean", "adm_med", "adm_75th", "adm_25th", "adm_975th", "adm_025th")
 
TN_adm_sum <- TN_adm_sum %>% as_tibble() %>% 
  mutate(obs = row_number(),
         Date = as.Date(start.date+obs))

```

### Diagnostics  
#### Traceplots 
```{r TN_traces}
traceplot(TN_fit,
          pars = names(stan_TN_adm)[which(grepl("duration", names(stan_TN_adm)))])
```

#### Posteriors  
```{r TN_posteriors}
stan_dens(TN_fit, 
          pars = names(stan_TN_adm)[which(grepl("duration", names(stan_TN_adm)))], 
          separate_chains = TRUE)

```

#### Comparison to observed data    

```{r TN_plot_fit}
ggplot() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11, 
                                   angle = 45,
                                   hjust = 1),
        axis.title = element_text(size = 14)) +
# Hospitalizations uncertainty intervals  
  geom_ribbon(data = TN_adm_sum,
              aes(x = Date, 
                  ymax = adm_75th,
                  ymin = adm_25th),
              fill = "darkblue",
              alpha = 0.3) +
  geom_ribbon(data = TN_adm_sum, 
              aes(x = Date, 
                  ymax = adm_975th,
                  ymin = adm_025th),
              fill = "darkblue",
              alpha = 0.1) +
# Observed data
  geom_point(data = TN_all,
             aes(x = as.Date(DATE), y = NEW_HOSP),
             pch = 17) +
# Model means
  geom_line(data = TN_adm_sum,
            aes(x = Date, y = adm_med),
            col = "darkblue") +
# Formatting
  scale_x_date(date_labels = "%m/%d", 
               date_breaks = "7 day",
               limits = as.Date(c("2020-02-16",
                                  "2020-07-01")))+
  ylim(c(0, max(StL$hosp.lower/0.9, na.rm = T)+50)) +
  labs(y = "Hospital Admissions",
       title = "TN hospital admissions fit")

```

